---
title: "analysisHQ"
author: "JD Pruett"
date: "2025-04-04"
---

# Environment Setup --------------------------------------
```{r setup, include=FALSE}
# knitr::opts_knit$set(root.dir = '/path/to/your/CES Accuracy Analysis')
knitr::opts_knit$set(root.dir = '/Users/jdpruett/Desktop/CES Accuracy Analysis')
```

```{r}
# Load functions and structures from other utility scripts
source("code/utilities/recoding_utilities.R")
source("code/utilities/data_structures.R")
source("code/utilities/election_return_utilities.R")
source("code/utilities/accuracy_utilities.R")
source("code/utilities/accuracy_utilities.R")
```

```{r, echo=False, warnings=FALSE, message=FALSE}
# create notin operator
`%notin%` <- Negate(`%in%`)

# Packages & working directory
pckgs <- c("haven", "readxl", "ipumsr", "here", "dplyr", "cdlTools", "anesrake", "writexl", "tidyverse", "rlang", "tibble", "naniar", "weights", "lubridate", "openxlsx", "cpsR", "purrr", "progress", "tidycensus", "humaniformat", "stringdist", "glue", "stringi")

## Uncomment this to download all necessary packages
# if (any(pckgs %notin% rownames(installed.packages())==TRUE)){
# install.packages(pckgs, repos = c(CRAN = "http://cloud.r-project.org"))}

invisible(sapply(pckgs, FUN = require, character.only = TRUE))

# current project root as identified by {here}
set_here()
```

# Read-in Data ----------------------------------------------
## CES Datasets
```{r}
CES_06_raw <- read_dta(here("data/CES Data/cces_2006_common.dta"))
CES_08_raw <- read_dta(here("data/CES Data/cces_2008_common.dta")) %>% 
  mutate(StateFIPS = as.numeric(V265),
        Congressional_District = na_if(V264, "") %>% 
        str_pad(width = 2, pad = "0"))
CES_10_raw <- read_dta(here("data/CES Data/cces_2010_common_validated.dta"))
CES_12_raw <- read_dta(here("data/CES Data/commoncontent2012.dta"))
CES_14_raw <- read_dta(here("data/CES Data/CCES14_Common_Content_Validated.dta"))
CES_16_raw <- read_dta(here("data/CES Data/CCES16_Common_OUTPUT_Feb2018_VV.dta"))
CES_18_raw <- read_dta(here("data/CES Data/cces18_common_vv.dta"))
CES_20_raw <- read_dta(here("data/CES Data/CES20_Common_OUTPUT_vv.dta"))
CES_22_raw <- read_dta(here("data/CES Data/CES_2022_Common_Output.dta"))
```

## Augment 2008 with Candidate File
The CES generally included candidate names as variables in the primary data release. In 2008, that file was released separately. Here, I load that file and add U.S. House and Senate candidate names to the 2008 CES df.
```{r}
ces_08_cand_name_file <- suppressWarnings(
  read_csv(
    "data/misc./2008 CCES Election File (House and Senate).csv",
    col_types = cols_only(
      `Election` = col_character(),
      `State FIPS Code` = col_integer(),
      District          = col_character(),
      `Democrat Name`   = col_character(),
      `Republican Name` = col_character(),
      `Other Name`      = col_character(),
      `Other name 2`    = col_character()
    ),
    show_col_types = FALSE
  )
) %>%
  rename(StateFIPS = `State FIPS Code`) %>%
  mutate(
    Congressional_District = str_pad(District, width = 2, pad = "0")
  )

# senate lookup
ces_08_us_senate_names <- ces_08_cand_name_file %>%
  filter(Election == "S") %>%
  select(
    StateFIPS,
    US_Senate_Democrat_Name   = `Democrat Name`,
    US_Senate_Republican_Name   = `Republican Name`,
    US_Senate_Other_Name = `Other Name`
  ) %>% 
  # Need to excude special election cases
  group_by(StateFIPS) %>%
  mutate(row_in_state = row_number()) %>%
  ungroup() %>%
  filter(!(StateFIPS %in% c(28, 56) & row_in_state == 2)) %>%
  select(-row_in_state)


# house lookup (state + district) 
ces_08_us_house_names <- ces_08_cand_name_file %>%
  filter(Election == "H") %>%
  select(
    StateFIPS,
    Congressional_District,
    US_House_Democrat_Name   = `Democrat Name`,
    US_House_Republican_Name   = `Republican Name`,
    US_House_Other_Name = `Other Name`
  )


# join everything in one pass
CES_08_raw <- CES_08_raw %>%
  left_join(ces_08_us_senate_names, by = "StateFIPS") %>%
  left_join(ces_08_us_house_names, by = c("StateFIPS", "Congressional_District"))
```

## Election Returns
```{r}
# President, Governor, U.S. Senate, U.S. House
President_0820 <- read_excel("data/Benchmarks/Election Returns/President_Returns_2008_2020.xlsx")
Governor_0622 <- read_excel("data/Benchmarks/Election Returns/Governor_Returns_2006_2022.xlsx")
Senate_0622 <- read_excel("data/Benchmarks/Election Returns/US_Senate_Returns_2006_2022.xlsx") %>% 
  filter(!(concurrent_special == TRUE & special == TRUE)) %>% 
  select(-concurrent_special) # concurrent_special == TRUE indicates a state/year with both a regular and a special Senate election.
# These cases are excluded because the CES typically asks about only one Senate race, 
# making it ambiguous which election respondents are reporting on.
# If there was only a single election (even if special), it would still be captured.
House_0622 <- read_excel("data/Benchmarks/Election Returns/US_House_Returns_2006_2022.xlsx", guess_max = 20000)

# State Senate, State House, Attorney General, Secretary of State
State_legislative_0623_raw <- read_excel("data/benchmarks/Election Returns/State_Legislative_and_Constitutional_Offices_2006_2023.xlsx", guess_max = 20000) %>% mutate(year = year(election_date))
# Ballotpedia mislabeled some elections as general elections that were actually special in their initial data query. This is a list of candidates that should be removed from State_legislative_0623 to avoid mixing in special elections.
special_candidates_to_remove <- read_excel("data/misc./mixed_in_special_election_candidates.xlsx")
State_legislative_0623 <- State_legislative_0623_raw %>%
  anti_join(special_candidates_to_remove,
            by = c("office_district", "candidate", "year")) %>%
  select(-year)

State_senate_full <- State_legislative_0623 %>%
  filter(office == "State Senate")
State_house_full <- State_legislative_0623 %>%
  filter(office == "State House")
Attorney_general_full <- State_legislative_0623 %>%
  filter(office == "Attorney General")
Secretary_of_state_full <- State_legislative_0623 %>%
  filter(office == "Secretary of State")
```

### Turnout
```{r}
turnout_national <- read_excel("data/Benchmarks/VAP_Turnout_Rates.xlsx", sheet="National")
turnout_statewide <- read_excel("data/Benchmarks/VAP_Turnout_Rates.xlsx", 
                                sheet = "Statewide", 
                                skip = 1, 
                                .name_repair = "unique_quiet") %>% 
  select(1:11) %>%
  slice(-n()) %>%
  mutate(State = toupper(State))
```

### State Populations
```{r}
state_populations <- read_excel("data/Benchmarks/Historical_State_Population_by_Year.xlsx")
```

### Misc.
```{r}
# CES assigned some respondents to incorrect congressional district in 2020
NC_flag_df <- read.csv("data/misc./NC_responents_incorrect_CD.csv") %>%
  filter(cd_change_flag_post == 1)

# This is a list of all the races that should be skipped because they were asked about in the CES but did not actually occur
skip_conditions <- read.xlsx("data/misc./skip_conditions_final.xlsx")
```

## CPS Datasets
```{r}
# Set IPUMS API Key (specific to user, free to acquire on IPUMS site)
USER_IPUMS_API_KEY = "59cba10d8a5da536fc06b59dd33fd321a4eb480fa8fe8ad46c9d0d3e"
set_ipums_api_key(USER_IPUMS_API_KEY, save = TRUE, overwrite = TRUE)
```

### Prep monthly sample codes and variables
This is necessary to use the IPUMS API to download CPS data.
```{r}
# All
cps_vars_all <- c("WTFINL", "EARNWT", "VOSUPPWT", "STATEFIP", "SEX", "EDUC", "AGE", "CITIZEN", "HISPAN", "OCC", "VETSTAT", "REGION", "EMPSTAT", "CLASSWKR", "UNION", "FAMINC", "VOTERES", "VOREG", "VOTED", "VOWHYNOT", "VOTEHOW")

# 2006
cps_sample_2006 = c('cps2006_01s', 'cps2006_02b', 'cps2006_03b', 'cps2006_04b', 'cps2006_05s', 'cps2006_06s', 'cps2006_07b', 'cps2006_08s', 'cps2006_09s', 'cps2006_10s', 'cps2006_11s', 'cps2006_12s')
# 2008
cps_sample_2008 = c('cps2008_01s', 'cps2008_02b', 'cps2008_03b', 'cps2008_04b', 'cps2008_05s', 'cps2008_06s', 'cps2008_07b', 'cps2008_08s', 'cps2008_09s', 'cps2008_10s', 'cps2008_11s', 'cps2008_12s')
# 2010
cps_sample_2010 = c('cps2010_01s', 'cps2010_02b', 'cps2010_03b', 'cps2010_04b','cps2010_05s', 'cps2010_06s','cps2010_07s', 'cps2010_08s', 'cps2010_09s', 'cps2010_10s', 'cps2010_11s', 'cps2010_12s')
# 2012
cps_sample_2012 = c('cps2012_01s', 'cps2012_02b', 'cps2012_03b', 'cps2012_04b', 'cps2012_05s', 'cps2012_06s', 'cps2012_07s', 'cps2012_08s', 'cps2012_09s', 'cps2012_10s', 'cps2012_11s', 'cps2012_12s')
# 2014
cps_sample_2014 = c('cps2014_01s', 'cps2014_02s', 'cps2014_03b', 'cps2014_04b', 'cps2014_05b', 'cps2014_06s', 'cps2014_07s', 'cps2014_08s', 'cps2014_09s', 'cps2014_10s', 'cps2014_11s', 'cps2014_12s')
# 2016
cps_sample_2016 = c('cps2016_01s', 'cps2016_02s', 'cps2016_03b', 'cps2016_04b', 'cps2016_05b', 'cps2016_06s', 'cps2016_07b', 'cps2016_08s', 'cps2016_09s', 'cps2016_10s', 'cps2016_11s', 'cps2016_12s')
# 2018
cps_sample_2018 = c('cps2018_01s', 'cps2018_02s', 'cps2018_03b', 'cps2018_04b', 'cps2018_05s', 'cps2018_06s', 'cps2018_07s', 'cps2018_08s', 'cps2018_09s', 'cps2018_10s', 'cps2018_11s', 'cps2018_12s')
# 2020
cps_sample_2020 = c('cps2020_01s', 'cps2020_02s', 'cps2020_03b', 'cps2020_04b', 'cps2020_05b', 'cps2020_06s', 'cps2020_07b', 'cps2020_08s', 'cps2020_09b', 'cps2020_10s', 'cps2020_11s', 'cps2020_12s')
# 2022
cps_sample_2022 = c('cps2022_01s', 'cps2022_02s', 'cps2022_03b', 'cps2022_04s', 'cps2022_05s', 'cps2022_06s', 'cps2022_07s', 'cps2022_08s', 'cps2022_09s', 'cps2022_10s', 'cps2022_11s', 'cps2022_12s')

```

### Read in CPS datasets to global environment
Note this creates xml (metadata) and .dat.giz (compressed raw data) files in the data/IPUMS CPS Artifacts subdirectory
```{r}
extract_codes <- list()

for (year in seq(2006, 2022, by = 2)) {
  suffix <- substr(year, 3, 4)
  extract_name <- paste("CPS Extract for", year)
  cps_sample_var <- paste("cps_sample", year, sep = "_")
  df_name <- paste("CPS", suffix, "raw", sep = "_")

  extract_code <- paste0(
    "  define_extract_micro(\n",
    "    collection = \"cps\",\n",
    "    description = \"", extract_name, "\",\n",
    "    samples = ", cps_sample_var, ",\n",
    "    variables = cps_vars_all) %>%\n",
    "  submit_extract() %>%\n",
    "  wait_for_extract() %>%\n",
    "  download_extract(download_dir = \"data/IPUMS CPS Artifacts\") %>%\n",
    "  read_ipums_micro()"
  )

  extract_codes[[df_name]] <- extract_code
}

for (df_name in names(extract_codes)) {
  assign(df_name, eval(parse(text = extract_codes[[df_name]])), envir = .GlobalEnv)
}
```



## Subset datasets --------------
Subset CES to only include relevant variables
```{r}
CES_06_subset_vars <- c(
  "v1000", "v1001", "v1002", "v1003", "v1004", "v2004", "v2018", "v2020", "v3081", "v3083", "v1006", "v2030", "v2073", "v2082", "v2033", "v2032", "v3004", "v4004", "v4006", "v4005", "v4014", "v4015", "v4013", "v4017", "v4018", "v4019", "v4020", "v4021", "v5009", "v5010", "v5011", "v5012", "v5005", "v5006", "v5007", "v5008", "v5001", "v5002", "v5003", "v5004")

CES_08_subset_vars <- c(
  "V100", "V201", "V206", "V207", "V208", "V250", "V251", "V202", "V205", "V204", "V213", "V252", "V253", "V258", "V257", "V247", "V259", "V264", "V501", "V502", "V553", "V555", "V601", "V602", "V603", "V612", "V653", "V655", "V657", "V672", "CC328_1", "CC328_3", "CC334", "CC410", "CC411", "CC411b", "CC412", "CC413", "CC414_2", "CC414_1", "V209", "CC329", "CC333", "V246", "V203", "CC403", "CC405", "CC404", "CC326b", "CC337", "CC337S2", "CC340", "CC338", "V618", "V619", "US_Senate_Democrat_Name", "US_Senate_Republican_Name", "US_Senate_Other_Name", "US_House_Democrat_Name", "US_House_Republican_Name", "US_House_Other_Name")

CES_10_subset_vars <- c(
  "V100", "V101", "V202_post", "V204_post", "V205_post", "V206", "V206_post", "V207", "V208", "V213", "V252", "V254", "V276", "V276_post", "V539_post", "V564_post", "V567_post", "V548_post", "V551_post", "V533_post", "V536_post", "V545_post", "V554_post", "V556_post", "V570_post", "V572_post","V574_post", "V576_post", "CC351", "V209", "V264", "V250", "V246", "V290", "CC401", "V203", "CC403", "CC402a", "CC410a", "CC410b", "CC412", "CC411", "CC413a", "CC413b", "CC413c", "CC413d")

CES_12_subset_vars <- c(
  "V101", "V103", "inputzip", "votereg_f", "regzip", "inputstate", "inputstate_post", "cdid113", "cdid113_post", "gender", "educ", "birthyr", "milstat_1", "milstat_3", "CC351", "votereg", "employ", "employercat", "union", "ownhome", "faminc",
  "hispanic", "CC401", "CC402a", "CC403", "CC410a", "CC410b", "CC355", "CC412", "CC411", "CC413a", "CC413b", "CC413c", "CC413d", "GovCand1Name_post", "GovCand1Party_post", "GovCand2Name_post", "GovCand2Party_post", "SenCand1Name_post", "SenCand1Party_post", "SenCand2Name_post", "SenCand2Party_post", "SenCand3Name_post", "SenCand3Party_post", "HouseCand1Name_post", "HouseCand1Party_post", "HouseCand2Name_post", "HouseCand2Party_post", "HouseCand3Name_post", "HouseCand3Party_post")


CES_14_subset_vars <- c(
  "V101", "weight", "inputzip_post", "votereg_f_post", "regzip_post", "inputstate", "inputstate_post", "cdid", "cdid_post", "gender", "educ", "birthyr", "milstat_1", "milstat_3", "CC351", 
  "votereg", "employ", "union", "ownhome", "faminc",
  "hispanic", "CC401", "CC402a", "CC403", "CC410b", "CC355x", "CC412", "CC411", "CC413a", "CC413b", "CC413c", "CC413d", "GovCand1Name_post", "GovCand1Party_post", "GovCand2Name_post", "GovCand2Party_post", "SenCand1Name_post", "SenCand1Party_post", "SenCand2Name_post", "SenCand2Party_post", "SenCand3Name_post", "SenCand3Party_post", "HouseCand1Name_post", "HouseCand1Party_post", "HouseCand2Name_post", "HouseCand2Party_post", "HouseCand3Name_post", "HouseCand3Party_post")

CES_16_subset_vars <- c(
  "V101", "inputzip_post", "votereg_f_post", "regzip_post", "commonweight_vv", "commonweight_vv_post", "commonweight_vv_lgbt", "commonweight", "commonweight_post", "inputstate", "inputstate_post", "cdid115", "cdid115_post", "gender", "educ", "birthyr", "milstat_1", "milstat_3", "CC16_361", "votereg", "employ", "union", "ownhome", "faminc", "hispanic", "CC16_401", "CC16_402a", "CC16_410a", "CC16_403", "CC16_410b", "CC16_412", "CC16_411", "CC16_413a", "CC16_413b", "CC16_413c", "CC16_413d", "GovCand1Name_post", "GovCand1Party_post", "GovCand2Name_post", "GovCand2Party_post", "GovCand3Name_post", "GovCand3Party_post", "SenCand1Name_post", "SenCand1Party_post", "SenCand2Name_post", "SenCand2Party_post", "SenCand3Name_post", "SenCand3Party_post", "SenCand4Name_post", "SenCand4Party_post", "HouseCand1Name_post", "HouseCand1Party_post", "HouseCand2Name_post", "HouseCand2Party_post", "HouseCand3Name_post", "HouseCand3Party_post", "HouseCand4Name_post", "HouseCand4Party_post", "HouseCand5Name_post", "HouseCand5Party_post", "HouseCand6Name_post", "HouseCand6Party_post", "HouseCand7Name_post", "HouseCand7Party_post", "HouseCand8Name_post", "HouseCand8Party_post", "HouseCand9Name_post", "HouseCand9Party_post", "HouseCand10Name_post", "HouseCand10Party_post", "HouseCand11Name_post", "HouseCand11Party_post")

CES_18_subset_vars <- c(
  "caseid", "inputzip_post", "votereg_f_post", "regzip_post", "commonweight", "commonpostweight", "vvweight", "vvweight_post", "inputstate", "inputstate_post", "cdid116", "cdid116_post", "gender", "educ", "birthyr", "cit1", "milstat_1", "milstat_3", "region", "votereg", "employ", "union", "ownhome", "faminc_new", "hispanic", "CC18_401", "CC18_402a", "CC18_403", "CC18_410b", "CC18_410bx", "CC18_412", "CC18_353x", "CC18_411", "CC18_352x", "CC18_420a", "CC18_420b", "CC18_420c", "CC18_413c", "CC18_413d", "GovCand1Name_post", "GovCand1Party_post", "GovCand2Name_post", "GovCand2Party_post", "GovCand3Name_post", "GovCand3Party_post", "SenCand1Name_post", "SenCand1Party_post", "SenCand2Name_post", "SenCand2Party_post", "SenCand3Name_post", "SenCand3Party_post", "AttCand1Name", "AttCand1Party", "AttCand2Name", "AttCand2Party", "SecCand1Name", "SecCand1Party", "SecCand2Name", "SecCand2Party", "HouseCand1Name_post", "HouseCand1Party_post", "HouseCand2Name_post", "HouseCand2Party_post", "HouseCand3Name_post", "HouseCand3Party_post", "HouseCand4Name_post", "HouseCand4Party_post", "HouseCand5Name_post", "HouseCand5Party_post", "HouseCand6Name_post", "HouseCand6Party_post", "HouseCand7Name_post", "HouseCand7Party_post")


CES_20_subset_vars <- c(
  "caseid", "inputzip_post", "votereg_f_post", "regzip_post", "commonweight", "commonpostweight", "vvweight", "vvweight_post", "inputstate", "inputstate_post", "cdid116", "cdid116_post", "gender", "educ", "birthyr", "cit1", "milstat_1", "milstat_3", "CC20_361", "region", "votereg", "employ", "union", "ownhome", "faminc_new", "hispanic", "CC20_401", "CC20_402a", "CC20_403", "CC20_410", "CC20_411", "CC20_365_voted", "CC20_411b", "CC20_365b_voted", "CC20_412", "CC20_413", "CC20_414a", "CC20_414b", "CC20_415c", "CC20_415d", "GovCand1Name_post", "GovCand1Party_post", "GovCand2Name_post", "GovCand2Party_post", "SenCand1Name_post", "SenCand1Party_post", "SenCand2Name_post", "SenCand2Party_post", "AttCand1Name", "AttCand1Party", "AttCand2Name", "AttCand2Party", "AttCand3Name", "AttCand3Party", "SecCand1Name", "SecCand1Party", "SecCand2Name", "SecCand2Party", "SecCand3Name", "SecCand3Party", "HouseCand1Name_post", "HouseCand1Party_post", "HouseCand2Name_post", "HouseCand2Party_post", "HouseCand3Name_post", "HouseCand3Party_post", "HouseCand4Name_post", "HouseCand4Party_post", "HouseCand5Name_post", "HouseCand5Party_post", "HouseCand6Name_post", "HouseCand6Party_post", "HouseCand7Name_post", "HouseCand7Party_post", "HouseCand8Name_post", "HouseCand8Party_post", "HouseCand9Name_post", "HouseCand9Party_post")

CES_22_subset_vars <- c(
"caseid", "inputzip_post", "votereg_f_post", "regzip_post", "commonweight", "commonpostweight", "inputstate", "inputstate_post", "cdid118", "cdid118_post", "gender4", "educ", "birthyr", "cit1", "milstat_1", "milstat_3", "CC22_361", "region", "votereg", "employ", "union", "ownhome", "faminc_new", "hispanic", "CC22_401", "CC22_402a", "CC22_403", "CC22_411", "CC22_365_voted", "CC22_411b", "CC22_365b_voted", "CC22_412", "CC22_413", "CC22_414a", "CC22_414b", "CC22_415c", "CC22_415d", "GovCand1Name_post", "GovCand1Party_post", "GovCand2Name_post", "GovCand2Party_post", "GovCand3Name_post", "GovCand3Party_post", "SenCand1Name_post", "SenCand1Party_post", "SenCand2Name_post", "SenCand2Party_post", "SenCand3Name_post", "SenCand3Party_post", "SenCand4Name_post", "SenCand4Party_post", "AttCand1Name", "AttCand1Party", "AttCand2Name", "AttCand2Party", "AttCand3Name", "AttCand3Party", "SecCand1Name", "SecCand1Party", "SecCand2Name", "SecCand2Party", "SecCand3Name", "SecCand3Party", "HouseCand1Name_post", "HouseCand1Party_post", "HouseCand2Name_post", "HouseCand2Party_post", "HouseCand3Name_post", "HouseCand3Party_post", "HouseCand4Name_post", "HouseCand4Party_post", "HouseCand5Name_post", "HouseCand5Party_post", "HouseCand6Name_post", "HouseCand6Party_post", "HouseCand7Name_post", "HouseCand7Party_post", "HouseCand8Name_post", "HouseCand8Party_post")

CES_06_raw_subset <- CES_06_raw[, CES_06_subset_vars]
CES_08_raw_subset <- CES_08_raw[, CES_08_subset_vars]
CES_10_raw_subset <- CES_10_raw[, CES_10_subset_vars]
CES_12_raw_subset <- CES_12_raw[, CES_12_subset_vars]
CES_14_raw_subset <- CES_14_raw[, CES_14_subset_vars]
CES_16_raw_subset <- CES_16_raw[, CES_16_subset_vars]
CES_18_raw_subset <- CES_18_raw[, CES_18_subset_vars]
CES_20_raw_subset <- CES_20_raw[, CES_20_subset_vars]
CES_22_raw_subset <- CES_22_raw[, CES_22_subset_vars]
```

Subset CPS to exclude respondents below 18 because the CES is limited to 18+
```{r}
# Define the years range
years <- seq(2006, 2022, by = 2)

# Iterate over the years
for (year in years) {
  # Extract the last two digits of the year
  year_short <- substr(year, 3, 4)
  
  # Define the variable names for the current year
  raw_df_name <- paste0("CPS_", year_short, "_raw")
  subset_vars <- intersect(colnames(get(raw_df_name)), cps_vars_all)

  # Subset the raw dataframe using the filtered variable list
  subset_df <- get(raw_df_name)[, subset_vars, drop = FALSE]
  
  # print any vars in cps_vars_all that are not in subset_vars
  print(paste("Missing vars for CES (0 expected)", year_short))
  print(setdiff(cps_vars_all, subset_vars))
  
  # Reorder the columns to match the order in cps_vars_all
  subset_df <- subset_df[, match(cps_vars_all, colnames(subset_df))]
  
  # Filter under 18
  subset_df <- subset_df %>%
    filter(AGE >= 18 & CITIZEN %in% c(1,2,3,4))  # Keep only citizens
  
  # Assign the subset dataframe back to the global environment
  assign(paste0(raw_df_name, "_subset"), subset_df, envir = .GlobalEnv)
}

```

### Update CES/CPS data structures
Sometimes I make changes to one year individually, and other times I make changes to all years simultaneously. To keep things clean, below are a series of functions to make sure that each individual year in the global environment is synced with each year in the "list_all_years" object, which groups all years together for easier processing.
```{r}
# list_all_years contains a list of all CES and CPS datasets, organized by year. As changes are made to these datasets, update_list_all_years updates list_all_years with the current versions of each df in the global environment.
update_list_all_years <- function() {
  list_all_years <- list(
  list(CPS = CPS_06_rc_clean, CES = CES_06_rc_clean, year = 2006),
  list(CPS = CPS_08_rc_clean, CES = CES_08_rc_clean, year = 2008),
  list(CPS = CPS_10_rc_clean, CES = CES_10_rc_clean, year = 2010),
  list(CPS = CPS_12_rc_clean, CES = CES_12_rc_clean, year = 2012),
  list(CPS = CPS_14_rc_clean, CES = CES_14_rc_clean, year = 2014),
  list(CPS = CPS_16_rc_clean, CES = CES_16_rc_clean, year = 2016),
  list(CPS = CPS_18_rc_clean, CES = CES_18_rc_clean, year = 2018),
  list(CPS = CPS_20_rc_clean, CES = CES_20_rc_clean, year = 2020),
  list(CPS = CPS_22_rc_clean, CES = CES_22_rc_clean, year = 2022)
  )
  
  assign("list_all_years", list_all_years, envir = .GlobalEnv)
}

# This does the inverse: If I make changes to each df inside of list_all_years, this function pushes those changes to the global environment, so each individual
# year reflects any changes made inside of list_all_years.
push_list_all_years_to_env <- function() {
  for (entry in list_all_years) {
    year <- entry$year
    
    # Construct object names
    cps_name <- paste0("CPS_", substr(year, 3, 4), "_rc_clean")
    ces_name <- paste0("CES_", substr(year, 3, 4), "_rc_clean")
    
    # Assign to global environment
    assign(cps_name, entry$CPS, envir = .GlobalEnv)
    assign(ces_name, entry$CES, envir = .GlobalEnv)
  }
}


update_ces_rc_datasets <- function() {
  ces_rc_datasets <- setNames(
  list(CES_06_rc, CES_08_rc, CES_10_rc, CES_12_rc, CES_14_rc, CES_16_rc, CES_18_rc, CES_20_rc, CES_22_rc),
  c("2006", "2008", "2010", "2012", "2014", "2016", "2018", "2020", "2022")
)
  assign("ces_rc_datasets", ces_rc_datasets, envir = .GlobalEnv)
}
```

# Recoding CES and CPS
## Codebook recodings
Recode CES and CPS to have corresponding levels for comparable variables. Ensure that CODEBOOK.xlsm is in the code directory at the same directory level as analysisHQ.Rmd. This function will output a recoded df in the form CES_YY_rc to the global environment. 

If alternate recodings are desired, make the change in the CODEBOOK, save it, and re-run the code below.
```{r message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE, results='hide'}
apply_codebook_and_recode(6)
apply_codebook_and_recode(8)
apply_codebook_and_recode(10)
apply_codebook_and_recode(12)
apply_codebook_and_recode(14)
apply_codebook_and_recode(16)
apply_codebook_and_recode(18)
apply_codebook_and_recode(20)
apply_codebook_and_recode(22)
```

## Manual recodings
```{r}
make_state_to_region <- function() {
  region_mapping <- list(
    "Northeast" = c("CT", "ME", "MA", "NH", "RI", "VT", "NJ", "NY", "PA"),
    "Midwest" = c("IL", "IN", "MI", "OH", "WI", "IA", "KS", "MN", "MO", "NE", "ND", "SD"),
    "South" = c("DE", "DC", "FL", "GA", "MD", "NC", "SC", "VA", "WV", "AL", "KY", "MS", "TN", "AR", "LA", "OK", "TX"),
    "West" = c("AZ", "CO", "ID", "MT", "NV", "NM", "UT", "WY", "AK", "CA", "HI", "OR", "WA"))

  # Flatten the mapping into a named vector
  state_abbr <- unlist(region_mapping)
  regions <- rep(names(region_mapping), times = lengths(region_mapping))
  setNames(regions, state_abbr)
}

state_to_region <- make_state_to_region()
```

```{r warning=FALSE}
year_mapping <- c("2006" = 1, "2008" = 2, "2010" = 3, "2012" = 4, "2014" = 5, "2016" = 6, "2018" = 7, "2020" = 8, "2022" = 9)

# Apply age groups to CES data frames
CES_06_rc$AGE_GROUP_rc <- calculate_age_groups(2006 - CES_06_rc$v2020)
CES_08_rc$AGE_GROUP_rc <- calculate_age_groups(2008 - CES_08_rc$V207)
CES_10_rc$AGE_GROUP_rc <- calculate_age_groups(2010 - CES_10_rc$V207)
CES_12_rc$AGE_GROUP_rc <- calculate_age_groups(2012 - CES_12_rc$birthyr)
CES_14_rc$AGE_GROUP_rc <- calculate_age_groups(2014 - CES_14_rc$birthyr)
CES_16_rc$AGE_GROUP_rc <- calculate_age_groups(2016 - CES_16_rc$birthyr)
CES_18_rc$AGE_GROUP_rc <- calculate_age_groups(2018 - CES_18_rc$birthyr)
CES_20_rc$AGE_GROUP_rc <- calculate_age_groups(2020 - CES_20_rc$birthyr)
CES_22_rc$AGE_GROUP_rc <- calculate_age_groups(2022 - CES_22_rc$birthyr)

# Apply age groups to CPS data frames
CPS_06_rc$AGE_GROUP_rc <- calculate_age_groups(CPS_06_rc$AGE)
CPS_08_rc$AGE_GROUP_rc <- calculate_age_groups(CPS_08_rc$AGE)
CPS_10_rc$AGE_GROUP_rc <- calculate_age_groups(CPS_10_rc$AGE)
CPS_12_rc$AGE_GROUP_rc <- calculate_age_groups(CPS_12_rc$AGE)
CPS_14_rc$AGE_GROUP_rc <- calculate_age_groups(CPS_14_rc$AGE)
CPS_16_rc$AGE_GROUP_rc <- calculate_age_groups(CPS_16_rc$AGE)
CPS_18_rc$AGE_GROUP_rc <- calculate_age_groups(CPS_18_rc$AGE)
CPS_20_rc$AGE_GROUP_rc <- calculate_age_groups(CPS_20_rc$AGE)
CPS_22_rc$AGE_GROUP_rc <- calculate_age_groups(CPS_22_rc$AGE)

# Add POST_STATE_rc to CPS data frames (have to include the 'post' so it aligns with CES)
data("fips_codes")
state_lookup <- fips_codes %>%
  select(state_code, state_name) %>%
  distinct() %>%
  mutate(
    state_code = as.integer(state_code),
    state_name = toupper(state_name)
  )

for (yr in c("06", "08", "10", "12", "14", "16", "18", "20", "22")) {
  df_name <- paste0("CPS_", yr, "_rc")
  df <- get(df_name)

  df <- df %>%
    mutate(STATEFIP = as.integer(STATEFIP)) %>%
    left_join(state_lookup, by = c("STATEFIP" = "state_code")) %>%
    rename(POST_STATE_rc = state_name)

  assign(df_name, df, envir = .GlobalEnv)
}

# Region
CES_06_rc$REGION_rc <- as.factor(state_to_region[CES_06_rc$v1002])
CES_08_rc$REGION_rc <- as.factor(state_to_region[fips(CES_08_rc$V206, to = "abbreviation")])
CES_10_rc$REGION_rc <- as.factor(state_to_region[fips(CES_10_rc$V206, to = "abbreviation")])
CES_12_rc$REGION_rc <- as.factor(state_to_region[fips(CES_12_rc$inputstate, to = "abbreviation")])
CES_14_rc$REGION_rc <- as.factor(state_to_region[fips(CES_14_rc$inputstate, to = "abbreviation")])
CES_16_rc$REGION_rc <- as.factor(state_to_region[fips(CES_16_rc$inputstate, to = "abbreviation")])
CES_18_rc$REGION_rc <- as.factor(state_to_region[fips(CES_18_rc$inputstate, to = "abbreviation")])
CES_20_rc$REGION_rc <- as.factor(state_to_region[fips(CES_20_rc$inputstate, to = "abbreviation")])
CES_22_rc$REGION_rc <- as.factor(state_to_region[fips(CES_22_rc$inputstate, to = "abbreviation")])

# Rename congressional district column
CES_06_rc$CDID_post_rc <- CES_06_rc$v1003
CES_08_rc$CDID_post_rc <- CES_08_rc$V264
CES_10_rc$CDID_post_rc <- CES_10_rc$V276_post
CES_12_rc$CDID_post_rc <- CES_12_rc$cdid113_post
CES_14_rc$CDID_post_rc <- CES_14_rc$cdid_post
CES_16_rc$CDID_post_rc <- CES_16_rc$cdid115_post
CES_18_rc$CDID_post_rc <- CES_18_rc$cdid116_post
CES_20_rc$CDID_post_rc <- CES_20_rc$cdid116_post
CES_22_rc$CDID_post_rc <- CES_22_rc$cdid118_post

# Add pre- and post- questionnaire state
update_ces_rc_datasets()
add_location_vars(ces_rc_datasets, pre_post_location_vars)
update_ces_rc_datasets()

# Ensure POST_STATE_rc is a factor
CES_06_rc$POST_STATE_rc <- as.factor(CES_06_rc$POST_STATE_rc)
CES_08_rc$POST_STATE_rc <- as.factor(CES_08_rc$POST_STATE_rc)
CES_10_rc$POST_STATE_rc <- as.factor(CES_10_rc$POST_STATE_rc)
CES_12_rc$POST_STATE_rc <- as.factor(CES_12_rc$POST_STATE_rc)
CES_14_rc$POST_STATE_rc <- as.factor(CES_14_rc$POST_STATE_rc)
CES_16_rc$POST_STATE_rc <- as.factor(CES_16_rc$POST_STATE_rc)
CES_18_rc$POST_STATE_rc <- as.factor(CES_18_rc$POST_STATE_rc)
CES_20_rc$POST_STATE_rc <- as.factor(CES_20_rc$POST_STATE_rc)
CES_22_rc$POST_STATE_rc <- as.factor(CES_22_rc$POST_STATE_rc)
                                     
                                     
cps_rc_datasets <- list(CPS_06_rc, CPS_08_rc, CPS_10_rc, CPS_12_rc, CPS_14_rc, CPS_16_rc, CPS_18_rc, CPS_20_rc, CPS_22_rc)

list2env(setNames(cps_rc_datasets, c("CPS_06_rc", "CPS_08_rc", "CPS_10_rc", "CPS_12_rc", "CPS_14_rc", "CPS_16_rc", "CPS_18_rc", "CPS_20_rc", "CPS_22_rc")), envir = .GlobalEnv)


# Clean up NAs (ie. make sure "NA", __NA__, "NA_char" are all treated as data type NA)
CES_06_rc_clean <- clean_NAs(CES_06_rc)
CPS_06_rc_clean <- clean_NAs(CPS_06_rc)

CES_08_rc_clean <- clean_NAs(CES_08_rc)
CPS_08_rc_clean <- clean_NAs(CPS_08_rc)

CES_10_rc_clean <- clean_NAs(CES_10_rc)
CPS_10_rc_clean <- clean_NAs(CPS_10_rc)

CES_12_rc_clean <- clean_NAs(CES_12_rc)
CPS_12_rc_clean <- clean_NAs(CPS_12_rc)

CES_14_rc_clean <- clean_NAs(CES_14_rc)
CPS_14_rc_clean <- clean_NAs(CPS_14_rc)

CES_16_rc_clean <- clean_NAs(CES_16_rc)
CPS_16_rc_clean <- clean_NAs(CPS_16_rc)

CES_18_rc_clean <- clean_NAs(CES_18_rc)
CPS_18_rc_clean <- clean_NAs(CPS_18_rc)

CES_20_rc_clean <- clean_NAs(CES_20_rc)
CPS_20_rc_clean <- clean_NAs(CPS_20_rc)

CES_22_rc_clean <- clean_NAs(CES_22_rc)
CPS_22_rc_clean <- clean_NAs(CPS_22_rc)

CES_06_rc_clean <- as.data.frame(CES_06_rc_clean)
CES_08_rc_clean <- as.data.frame(CES_08_rc_clean)
CES_10_rc_clean <- as.data.frame(CES_10_rc_clean)
CES_12_rc_clean <- as.data.frame(CES_12_rc_clean)
CES_14_rc_clean <- as.data.frame(CES_14_rc_clean)
CES_16_rc_clean <- as.data.frame(CES_16_rc_clean)
CES_18_rc_clean <- as.data.frame(CES_18_rc_clean)
CES_20_rc_clean <- as.data.frame(CES_20_rc_clean)
CES_22_rc_clean <- as.data.frame(CES_22_rc_clean)

update_list_all_years()
```

# TEMP
```{r}
CES_06_rc_clean <- clean_NAs(CES_06_rc)
CES_08_rc_clean <- clean_NAs(CES_08_rc)
CES_10_rc_clean <- clean_NAs(CES_10_rc)
CES_12_rc_clean <- clean_NAs(CES_12_rc)
CES_14_rc_clean <- clean_NAs(CES_14_rc)
CES_16_rc_clean <- clean_NAs(CES_16_rc)
CES_18_rc_clean <- clean_NAs(CES_18_rc)
CES_20_rc_clean <- clean_NAs(CES_20_rc)
CES_22_rc_clean <- clean_NAs(CES_22_rc)

CES_06_rc_clean <- as.data.frame(CES_06_rc_clean)
CES_08_rc_clean <- as.data.frame(CES_08_rc_clean)
CES_10_rc_clean <- as.data.frame(CES_10_rc_clean)
CES_12_rc_clean <- as.data.frame(CES_12_rc_clean)
CES_14_rc_clean <- as.data.frame(CES_14_rc_clean)
CES_16_rc_clean <- as.data.frame(CES_16_rc_clean)
CES_18_rc_clean <- as.data.frame(CES_18_rc_clean)
CES_20_rc_clean <- as.data.frame(CES_20_rc_clean)
CES_22_rc_clean <- as.data.frame(CES_22_rc_clean)

update_list_all_years()
```

# RECODE PARTY AND CANDIDATE VOTE
For 2006 and 2008, respondent vote choices could be directly mapped to a party because the response options were simply "Democrat," "Republican," and "Other". However, from 2010 - 2022, the CES changed the response options. Instead of using party names, respondents were provided candidate names. These names were inputted dynamically based for a particular respondent based on their state of residence.

This means that error can be calculated in two ways: at the party level (by aggregating across all candidates of a particular party), and at the candidate level (by comparing CES votes for a particular candidate to actual votes received by that specific candidate).

Thus, I create two variables for each CES respondent. The first will be named [RACE]_PARTY_rc and will contain indicate the Party that the respondent voted for. In the cases when a respondent was presented a candidate name, the party will be filled in based on the provided contextual variables. The second will be [RACE]_CANDIDATE_rc and will contain the name of the candidate that the respondent voted for.

Note: Once you have run the two functions below once, you cannot run them again on the same dfs. You would need to go back and re-run # Clean up NAs.
```{r}
# The CES raw data contains two cases where a candidate's party was omitted.
# These are manually corrected here before running the recoding functions below.
# (1) Nick Begich III (R), AK US House, 2022
# (2) Sam Petsonk (D), WV Attorney General, 2020

# Nick Begich III - AK, US House, 2022 - missing party
CES_22_rc_clean <- CES_22_rc_clean %>%
  mutate(HouseCand3Party_post = if_else(
    HouseCand3Name_post == "Nick Begich III" & POST_STATE_rc == "ALASKA",
    "Republican",
    HouseCand3Party_post
  ))

# Sam Petsonk - WV, AG, 2020 - missing party
CES_20_rc_clean <- CES_20_rc_clean %>%
  mutate(AttCand1Party = if_else(
    AttCand1Name == "Sam Petsonk" & POST_STATE_rc == "WEST VIRGINIA",
    "Democrat",
    AttCand1Party
  ))

# (3) Joe Dorman (D) and Mary Fallin (R), OK Governor, 2014
# Some rows were missing both candidates' names and parties for this race.
CES_14_rc_clean <- CES_14_rc_clean %>%
  mutate(
    GovCand1Name_post = if_else(POST_STATE_rc == "OKLAHOMA" & (is.na(GovCand1Name_post) | GovCand1Name_post == ""), "Joe Dorman", GovCand1Name_post),
    GovCand1Party_post = if_else(POST_STATE_rc == "OKLAHOMA" & (is.na(GovCand1Party_post) | GovCand1Party_post == ""), "Democratic", GovCand1Party_post),
    GovCand2Name_post = if_else(POST_STATE_rc == "OKLAHOMA" & (is.na(GovCand2Name_post) | GovCand2Name_post == ""), "Mary Fallin", GovCand2Name_post),
    GovCand2Party_post = if_else(POST_STATE_rc == "OKLAHOMA" & (is.na(GovCand2Party_post) | GovCand2Party_post == ""), "Republican", GovCand2Party_post)
  )

update_list_all_years()
```

```{r}
#' Recode Party and Candidate Vote Variables from Dynamic References
#'
#' If a cell is "Democrat", "Republican", "Other", or NA, it’s left alone. Otherwise, the cell’s 
#' value is treated as a column name (e.g., "PRES_PARTY_post"). If that column exists, the cell is 
#' replaced with the value from that referenced column on the same row.

#' @return A list of the same structure, with each `CES` data.frame updated
#'   in-place so that dynamic references (column names stored in cells)
#'   have been replaced by their actual values.
#'
#' @details  
#' - Columns matching `"_PARTY_rc$"` (except those containing `"SENATOR2"` or
#'   `"SENATOREARLY"`) are always recoded.  
#' - Columns matching `"_CANDIDATE_rc$"` (same exclusions) are recoded for
#'   years **other than** 2006 and 2008.  
#' - Factor columns in each `CES` frame are coerced to character prior to
#'   recoding, and the recoded party columns are converted back to factors.
#'
recode_party_and_candidate_votes <- function(list_all_years) {
  no_recode <- c("Democrat", "Republican", "Other")  # values to be left alone
  
  n_years <- length(list_all_years)
  pb <- txtProgressBar(min = 0, max = n_years, style = 3)

  for (i in seq_along(list_all_years)) {
    year <- list_all_years[[i]]$year
    setTxtProgressBar(pb, i)
    cat(sprintf("  Processing year: %s\r", year))
    flush.console()

    year_entry <- list_all_years[[i]]
    year_chr   <- as.character(year_entry$year)
    df         <- year_entry$CES

    # identify the columns to fix
    party_cols     <- grep("_PARTY_rc$", names(df), value = TRUE)
    party_cols     <- party_cols[!grepl("SENATOR2|SENATOREARLY", party_cols)]
    candidate_cols <- grep("_CANDIDATE_rc$", names(df), value = TRUE)
    candidate_cols     <- candidate_cols[!grepl("PRES|SENATOR2|SENATOREARLY", candidate_cols)]
    
    cols_to_edit <- unique(c(party_cols, candidate_cols))
    df[cols_to_edit] <- lapply(df[cols_to_edit], as.character)


    # --- PARTY columns ---
    for (col in party_cols) {
      for (r in seq_len(nrow(df))) {
        val <- df[[col]][r]
        if (!(val %in% no_recode) && !is.na(val)) {
          # val should be the name of a contextual variable column
          if (! val %in% names(df)) {
            stop(sprintf(
              "Year %s: column '%s', row %d: reference '%s' not found",
              year_chr, col, r, val
            ))
          }
          df[[col]][r] <- as.character(df[[val]][r])
        }
      }
    }
    
    # # --- CANDIDATE columns (skip 2006 & 2008) ---
    # if (! year_chr %in% c("2006")) {
    #   for (col in candidate_cols) {
    #     for (r in seq_len(nrow(df))) {
    #       val <- df[[col]][r]
    #       if (!(val %in% no_recode) && !is.na(val)) {
    #         if (! val %in% names(df)) {
    #           stop(sprintf(
    #             "Year %s: column '%s', row %d: reference '%s' not found",
    #             year_chr, col, r, val
    #           ))
    #         }
    #         df[[col]][r] <- as.character(df[[ val ]][r])
    #       }
    #     }
    #   }
    # }

    # --- CANDIDATE columns (skip 2006 & 2008) ---
    if (! year_chr %in% c("2006")) {
      for (col in candidate_cols) {
        for (r in seq_len(nrow(df))) {
          val <- df[[col]][r]
          if (!(val %in% no_recode) && !is.na(val)) {
            if (! val %in% names(df)) {
              stop(sprintf(
                "Year %s: column '%s', row %d: reference '%s' not found",
                year_chr, col, r, val
              ))
            }
            # In some cases, even though a respondent selected the "Other"
            # candidate, no name is provided for that candidate. In those cases,
            # fill in manually.
            referenced_value <- as.character(df[[val]][r])
            
            # Derive corresponding party column
            party_col <- sub("_CANDIDATE_rc$", "_PARTY_rc", col)
            party_val <- if (party_col %in% names(df)) as.character(df[[party_col]][r]) else NA
            
            if (is.na(referenced_value) || referenced_value == "") {
              # Only fill placeholder if party is not blank/NA
              if (!is.na(party_val) && party_val != "") {
                df[[col]][r] <- "Other Candidate - No Name Provided"
              }
              # Otherwise leave it as NA (do nothing)
            } else {
              df[[col]][r] <- referenced_value
            }
          }
        }
      }
    }
    
    df[party_cols] <- lapply(df[party_cols], factor)

    # put the modified CES DF back
    list_all_years[[i]]$CES <- df
  }

  close(pb)
  list_all_years
}


```

Function to clean all the PARTY columns, to ensure consistent naming
```{r}
#' Standardize Party Labels in CES Data
#'
#' This function iterates over each year's CES data in a list and standardizes the values
#' in all `*_PARTY_rc` columns according to a set of cleaning rules:
#'
#' - `"Democrat"` and `"Republican"` are kept as-is.
#' - `"Democratic-Republican"` is recoded as `"Other"`.
#' - Any other value containing `"Democratic"` is recoded to `"Democrat"`.
#' - All other values (e.g., `"Green"`, `"Libertarian"`, NA) are recoded to `"Other"`.
#'
#' After recoding, all `*_PARTY_rc` columns are re-cast as factors with levels
#' `c("Democrat", "Republican", "Other")`.
#'
#' @param list_all_years A named list of lists, each containing a CES data frame under `$CES`.
#' @return The same list with party names cleaned in all `_PARTY_rc` columns.
#' @export
clean_party_names <- function(list_all_years) {
  for (i in seq_along(list_all_years)) {
    entry <- list_all_years[[i]]
    df    <- entry$CES
    party_cols <- grep("_PARTY_rc$", names(df), value = TRUE)

    for (col in party_cols) {
      original <- as.character(df[[col]])
      cleaned <- case_when(
        is.na(original) ~ NA,
        original == "Democrat" ~ "Democrat",
        original == "Republican" ~ "Republican",
        original == "Democratic-Republican" ~ "Other",
        grepl("Democratic", original, ignore.case = FALSE) ~ "Democrat",
        TRUE ~ "Other"
      )
      df[[col]] <- factor(cleaned, levels = c("Democrat", "Republican", "Other"))
    }

    list_all_years[[i]]$CES <- df
  }

  return(list_all_years)
}
```

Update list_all_years, run the recoding, clean party names, then push changes from list_all_years back to global environment to keep list_all_years synced with each individual CES year in the global environment.
```{r}
update_list_all_years()
recoded_party_candidate_list_all_years <- recode_party_and_candidate_votes(list_all_years)
recoded_party_candidate_list_all_years_cleaned <- clean_party_names(recoded_party_candidate_list_all_years)
rm(recoded_party_candidate_list_all_years)

list_all_years <- recoded_party_candidate_list_all_years_cleaned
push_list_all_years_to_env()
```

# ELECTION RETURN PROP TABLES
## Create Party-level Election Return Prop Tables
The raw election return files are loaded in the "Election Return Datasets" section. The following code creates a series of "prop" (proportion) tables for each election, then collapses across party to indicate what share of votes each party received in each election. 
```{r}
# US House
House_0622_clean_fixed <- recalculate_total_votes(House_0622, "statewide")
house_prop_table <- calculate_state_level_party_proportions(House_0622_clean_fixed) 
house_prop_table$Office <- "US House"

# US Senate
Senate_0622_clean_fixed <- recalculate_total_votes(Senate_0622, "statewide")
senate_cand_prop_table <- calculate_party_proportions(Senate_0622_clean_fixed) %>%
  mutate(office = "US Senate") %>%
  rename(
    Office = office,
    State = state,
    District = district,
    Year = year,
    #Candidate = candidate_name,
    Party_Detailed = party_detailed,
    Party_Simplified = party_simplified,
    #Candidate_Votes = candidatevotes,
    Total_Votes = totalvotes,
    Proportion = proportion
  )

# Governor
Governor_0622_clean_fixed <- recalculate_total_votes(Governor_0622, "statewide")
governor_cand_prop_table <- calculate_cand_num_proportions(Governor_0622_clean_fixed) %>%
  mutate(office = "Governor") %>%
  rename(
    Office = office,
    State = state,
    District = district,
    Year = year,
    Candidate = candidate_name,
    Cand_Num = cand_num,
    Party_Detailed = party_detailed,
    Party_Simplified = party_simplified,
    Candidate_Votes = candidatevotes,
    Total_Votes = totalvotes,
    Proportion = proportion
  )

# President
President_0820_nums <- assign_president_candidate_numbers(President_0820) %>%
  filter(candidate %notin% c("OVER VOTE", "OVERVOTE", "OVERVOTES", "OVER VOTES"))
president_candid_prop_table <- calculate_cand_num_proportions(President_0820_nums) %>%
  mutate(office = "President") %>%
  rename(
    Office = office,
    State = state,
    District = district,
    Year = year,
    Candidate = candidate_name,
    Cand_Num = cand_num,
    Party_Detailed = party_detailed,
    Party_Simplified = party_simplified,
    Candidate_Votes = candidatevotes,
    Total_Votes = totalvotes,
    Proportion = proportion
  )

# State Senate
state_senate_prop_table <- make_bp_prop_table(State_senate_full) %>%
  mutate(Office = "State Senator") %>%   
  select(Office, everything())    

# State House
state_house_prop_table <- make_bp_prop_table(State_house_full) %>%
  mutate(Office = "State Representative") %>%   
  select(Office, everything())    

# Attorney General
attorney_general_prop_table <- make_bp_prop_table(Attorney_general_full) %>%
  mutate(Office = "Attorney General") %>%   
  select(Office, everything()) 

# Secretary of State
secretary_of_state_prop_table <- make_bp_prop_table(Secretary_of_state_full) %>%
  mutate(Office = "Secretary of State") %>%   
  select(Office, everything()) 

merge_list <- list(
  governor_cand_prop_table,
  senate_cand_prop_table,
  state_senate_prop_table,
  state_house_prop_table,
  attorney_general_prop_table,
  secretary_of_state_prop_table,
  president_candid_prop_table
)

merged_prop_tables <- merge_list %>%
  map(~ .x %>%
        mutate(Year = as.integer(parse_number(as.character(Year)))) %>%
        # Candidate_Votes gets exclude because collapsing across candidates listed under multiple parties
        select(-any_of("Candidate_Votes")) 
      ) %>%
  bind_rows()
```

Collapsing prop table across party
```{r}
collapse_proportion_table <- function(df) {
  df %>%
    group_by(State, Year = as.numeric(Year), Office, Party_Simplified) %>%
    summarise(Proportion = sum(Proportion, na.rm = TRUE), .groups = "drop") %>%
    pivot_wider(
      names_from = Party_Simplified,
      values_from = Proportion,
      names_prefix = "",
      values_fill = 0
    ) %>%
    rename(
      DEM_Proportion = DEM,
      REP_Proportion = REP,
      OTHER_Proportion = OTHER
    ) %>%
    select(State, Year, Office, DEM_Proportion, OTHER_Proportion, REP_Proportion)
}

collapsed_merged_prop_table <- collapse_proportion_table(merged_prop_tables)
merged_prop_tables_combined <- rbind(collapsed_merged_prop_table, house_prop_table)
```


## Create Candidate-Level Election Return Prop Tables
Now I create a candidate-level prop table, which will be used to compare the CES vote to the actual votes received by each candidate in each election. This is done by creating a table that contains the proportion of votes received by each candidate in each state, for each year, and not grouping by party. 
```{r}
candidate_level_proportions <- function(df) {

  # summarize votes per candidate per race in case same candidate appears multiple times under different parties
  summed_votes <- df %>%
    group_by(year, state, office, district, candidate) %>%
    summarise(
      Candidate_Votes = sum(candidatevotes, na.rm = TRUE),
      Total_Votes = first(totalvotes),
      .groups = "drop"
    )

  # get party info from row with max candidate votes per group
  top_party <- df %>%
    group_by(year, state, office, district, candidate) %>%
    slice_max(candidatevotes, n = 1, with_ties = FALSE) %>%
    select(
      year, state, office, district, candidate,
      Party_Detailed = party_detailed,
      Party_Simplified = party_simplified
    )

  # combine and compute proportion
  cleaned <- summed_votes %>%
    left_join(top_party, by = c("year", "state", "office", "district", "candidate")) %>%
    mutate(Proportion = Candidate_Votes / Total_Votes) %>%
    select(
      Office = office,
      State = state,
      District = district,
      Year = year,
      Candidate = candidate,
      Party_Detailed,
      Party_Simplified,
      Candidate_Votes,
      Total_Votes,
      Proportion
    )

  return(cleaned)
}
```

```{r}
# US House
House_0622_district <- recalculate_total_votes(House_0622, "district")
house_candidate_level_props <- candidate_level_proportions(House_0622_district)
house_candidate_level_props$Office <- "US House"

# US Senate
senate_candidate_level_props  <- candidate_level_proportions(Senate_0622_clean_fixed)
senate_candidate_level_props$Office <- "US Senate"

# Governor
governor_candidate_level_props  <- candidate_level_proportions(Governor_0622_clean_fixed)
governor_candidate_level_props$Office <- "Governor"

# President
President_0820_candidate_level_props <- assign_president_candidate_numbers(President_0820) %>%
  filter(candidate %notin% c("OVER VOTE", "OVERVOTE", "OVERVOTES", "OVER VOTES"))

president_candidate_level_props <- calculate_cand_num_proportions(President_0820_candidate_level_props) %>%
  mutate(office = "President") %>%
  rename(
    Office = office,
    State = state,
    District = district,
    Year = year,
    Candidate = candidate_name,
    Party_Detailed = party_detailed,
    Party_Simplified = party_simplified,
    Candidate_Votes = candidatevotes,
    Total_Votes = totalvotes,
    Proportion = proportion
  ) %>%
  select(Office, State, District, Year, Candidate, Party_Detailed, Party_Simplified, Candidate_Votes, Total_Votes, Proportion)

# Attorney General
attorney_general_candidate_level_props <- candidate_level_BP_statewide_offices(Attorney_general_full) %>%
  mutate(Office = "Attorney General") %>%   
  select(Office, everything()) 

# Secretary of State
secretary_of_state_candidate_level_props <- candidate_level_BP_statewide_offices(Secretary_of_state_full) %>%
  mutate(Office = "Secretary of State") %>%   
  select(Office, everything()) 

# Merge all results
merged_candidate_level_props <- rbind(
    president_candidate_level_props,
    senate_candidate_level_props,
    house_candidate_level_props,
    governor_candidate_level_props,
    attorney_general_candidate_level_props,
    secretary_of_state_candidate_level_props
  ) %>%
  mutate(
    Candidate = Candidate %>%
      fix_suffix_position() %>%
      humaniformat::format_reverse(),
    State = toupper(State)
  )
```


# CREATE WEIGHTING TARGETS
ANESRake requires that "targets" are created for each variable, which specify the expected distribution of that variable in the population for each level (category) of that variable.
```{r}
# Some variables require special weighting variables. This dictionary specifies which weighting variable should be used for those special variables, otherwise
# the common weighting var is used. 
create_targets <- function(year_data_list) {
  cps_wtvar_dict <- c(
  "UNION_rc" = "EARNWT",
  "VOTED_rc" = "VOSUPPWT",
  "VOTEHOW_rc" = "VOSUPPWT",
  "VOTERES_rc" = "VOSUPPWT",
  "VOTEREG_rc" = "VOSUPPWT"
)
  
  # Create a list to store targets for each year
  all_targets <- list()
  
  # Loop through each year's data
  for (year_data in year_data_list) {
    cps_df <- year_data$CPS
    ces_df <- year_data$CES
    current_year <- year_data$year
    
    # Create a list of all recoded variables for CES data
    variables <- intersect(names(ces_df), names(cps_df))
    # Exclude variables I don't want to use in weighting
    variables <- variables[!grepl("DUM|OVERTIME|REGION_rc|CUR_MIL_rc|CLASSWKR_rc|CITIZEN_rc|POST_STATE|VOTEREG_rc", variables)]
  
    # Create population (cps) numeric values list with level names
    targets <- list()
    for (var in variables) {
      #browser()
      # We use actual state turnout figures, not CPS data, for VOTED_rc
      if (var == "VOTED_rc") {
        rate <- turnout_national %>%
          filter(Year == current_year) %>%
          pull(Rate)
        
        numeric_values <- c(rate, 1 - rate)
        names(numeric_values) <- c("Voted", "Did not vote")
        
      }
      
      else {
      # Determine the appropriate weight variable
      weight_var <- if (var %in% names(cps_wtvar_dict)) cps_wtvar_dict[var] else "WTFINL"
      
      # Calculate weighted proportions for each level using wpct
      weighted_proportions <- wpct(cps_df[[var]], weight = cps_df[[weight_var]], na.rm=TRUE)
      
      # Convert proportions to numeric values with level names
      numeric_values <- as.numeric(weighted_proportions)
      names(numeric_values) <- names(weighted_proportions)
      }

      # Store numeric values in the list
      targets[[var]] <- numeric_values
    }
    
    # Append targets for the current year to the list of all targets
    target_name <- paste0(current_year)
    all_targets[[target_name]] <- targets
  }
  
  return(all_targets)
}
```

Update list_all_years, then create targets for each year for all variables that are present in both CES and CPS.
```{r}
update_list_all_years()
all_targets <- create_targets(list_all_years)
```

# CREATES ANESRAKE WEIGHTS
I construct two sets of post-stratification weights with anesrake.

(1) anesrake_weight_full: An unrestricted weighting specification, in which all 
     available auxiliary variables are supplied to ANESRake. In this set, ANESRake 
     is free to select any number of variables for raking based on imbalance, and in
     practice it typically incorporates most of the available variables.

(2) anesrake_weight_restricted: A restricted specification, in which the set of 
     auxiliary variables is deliberately limited. This approach preserves a 
     larger set of variables for subsequent error evaluation, and only 
     supplies the most "discrepant" variables between CES and the targets
     to the ANESRake algorithm. 
     
## anesrake_weight_full
```{r}
run_weighting <- function(year_list,
                          targets,
                          year_data_list = list_all_years,
                          type = "pctlim",        # "pctlim", "nlim", "nmin", "nmax", or "nolim"
                          pctlim = 5,             # % threshold for "pctlim"
                          nlim = 5,               # n for "nlim"/"nmax"
                          choosemethod = "total", # how anesrakefinder scores discrepancy
                          verbose = FALSE) {
  results_list <- list()

  for (year in year_list) {
    message("Year: ", year)

    # Find the matching year data directly
    year_data <- year_data_list[[which(sapply(year_data_list, function(x) x$year == year))]]

    if (is.null(year_data)) {
      stop(glue::glue("Year {year} not found in year_data_list"))
    }

    target_data <- targets[[as.character(year)]]
    CES_data    <- year_data$CES
    caseid_col  <- CES_data[, 1]

    outsave <- anesrake(
      inputter     = target_data,
      dataframe    = CES_data,
      caseid       = caseid_col,
      verbose      = verbose,
      type         = type,
      pctlim       = pctlim,
      nlim         = nlim,
      choosemethod = choosemethod
    )

    caseweights <- data.frame(cases = outsave$caseid, weights = outsave$weightvec)

    # Which variables were actually used
    vars_used <- outsave$varsused

    # Make a tidy record of targets used (only those actually raked on)
    targets_used <- lapply(vars_used, function(v) {
      tv <- outsave$targets[[v]]
      data.frame(Variable = v,
                 Level    = names(tv),
                 Target   = as.numeric(tv),
                 row.names = NULL,
                 check.names = FALSE)
    })
    targets_used <- do.call(rbind, targets_used)

    # Store results
    results_list[[as.character(year)]] <- list(
      caseweights      = caseweights,
      vars_used        = vars_used,
      weighting_meta   = list(
        type          = outsave$type,
        choosemethod  = outsave$choosemethod,
        iterations    = outsave$iterations,
        converged     = outsave$converge,
        nonconvergence= outsave$nonconvergence
      ),
      targets_used     = targets_used,
      outsave_summary  = summary(outsave)  # pretty console summary
    )
  }

  return(results_list)
}
```

Create weights for all years
```{r}
weighting_results_all <- run_weighting(c(2006, 2008, 2010, 2012, 2014, 2016, 2018, 2020, 2022), all_targets)
```


Add the new weight column to each year's df
```{r}
add_weights_to_df <- function(df, year, weighting_results, col_name) {

  # Notify and remove existing weight column if present
  if (col_name %in% colnames(df)) {
    message(paste0(
      col_name,
      " already present in ",
      year,
      " — replacing with new weights."
    ))
    df <- df %>% dplyr::select(-all_of(col_name))
  }

  # Extract weights for the specified year
  weights <- weighting_results[[as.character(year)]]$caseweights
  weights <- weights[order(weights$cases), ]  # ensure consistent order

  # Merge weights with the dataframe
  merged_df <- merge(
    df,
    weights,
    by.x = 1,          # assumes case ID is first column of df
    by.y = "cases",
    all.x = TRUE
  )

  # Rename the weights column
  colnames(merged_df)[ncol(merged_df)] <- col_name

  merged_df
}

```

Add new weighting column to each CES df
```{r}
CES_06_rc_clean <- add_weights_to_df(CES_06_rc_clean, 2006, weighting_results_all, "anesrake_weight_full")
CES_08_rc_clean <- add_weights_to_df(CES_08_rc_clean, 2008, weighting_results_all, "anesrake_weight_full")
CES_10_rc_clean <- add_weights_to_df(CES_10_rc_clean, 2010, weighting_results_all, "anesrake_weight_full")
CES_12_rc_clean <- add_weights_to_df(CES_12_rc_clean, 2012, weighting_results_all, "anesrake_weight_full")
CES_14_rc_clean <- add_weights_to_df(CES_14_rc_clean, 2014, weighting_results_all, "anesrake_weight_full")
CES_16_rc_clean <- add_weights_to_df(CES_16_rc_clean, 2016, weighting_results_all, "anesrake_weight_full")
CES_18_rc_clean <- add_weights_to_df(CES_18_rc_clean, 2018, weighting_results_all, "anesrake_weight_full")
CES_20_rc_clean <- add_weights_to_df(CES_20_rc_clean, 2020, weighting_results_all, "anesrake_weight_full")
CES_22_rc_clean <- add_weights_to_df(CES_22_rc_clean, 2022, weighting_results_all, "anesrake_weight_full")
```
Table shows which variables were selected by anesrake for weighting in each year
```{r}
# ------------------------------------------------------------
# 1) Define the universe of variables across all years
# ------------------------------------------------------------
all_vars <- sort(unique(unlist(lapply(all_targets, names))))

# ------------------------------------------------------------
# 2) Build long table with 3-level status
# ------------------------------------------------------------
vars_status_long <- do.call(
  rbind,
  lapply(names(weighting_results_all), function(y_chr) {

    y <- as.character(y_chr)

    proposed <- names(all_targets[[y]])
    used     <- weighting_results_all[[y]]$vars_used

    data.frame(
      Year     = y,
      Variable = all_vars,
      Status   = dplyr::case_when(
        all_vars %in% used     ~ "Used",
        all_vars %in% proposed ~ "Provided but not selected",
        TRUE                   ~ "Not provided"
      ),
      stringsAsFactors = FALSE
    )
  })
)

# ------------------------------------------------------------
# 3) Pivot to wide (years as columns)
# ------------------------------------------------------------
vars_used_by_anesrake <- vars_status_long %>%
  pivot_wider(
    names_from  = Year,
    values_from = Status,
    values_fill = "Not provided"
  ) %>%
  arrange(Variable) %>%
  standardize_variable_names()
```

## anesrake_weight_restricted
First, I identify the variables that are consistently most discrepant from population benchmarks.
These variables are retained for ANESRake, as they capture the most "information" for post-stratification. 

```{r}
compute_year_discrepancies <- function(year,
                                       targets,
                                       year_data_list,
                                       choosemethod = "total") {
  
  # Get CES data for year
  year_data <- year_data_list[[which(sapply(year_data_list, function(x) x$year == year))]]
  CES_data  <- year_data$CES
  
  # All candidate targets for that year
  year_targets <- targets[[as.character(year)]]
  
  # Run anesrakefinder ONCE for the year
  finder <- anesrakefinder(
    inputter     = year_targets,
    dataframe    = CES_data,
    choosemethod = choosemethod
  )
  
  # Convert named vector → tidy df
  data.frame(
    Year        = year,
    Variable    = names(finder),
    Discrepancy = as.numeric(finder),
    stringsAsFactors = FALSE
  ) %>%
    arrange(desc(Discrepancy))
}

discrepancy_long <- map_dfr(
  c(2006, 2008, 2010, 2012, 2014, 2016, 2018, 2020, 2022),
  compute_year_discrepancies,
  targets = all_targets,
  year_data_list = list_all_years,
  choosemethod = "total"   # or "max", "average", etc.
) %>%
  group_by(Year) %>%
  mutate(
    Rank = dense_rank(desc(Discrepancy))
  ) %>%
  ungroup()

discrepancy_consistency <- discrepancy_long %>%
  group_by(Variable) %>%
  summarise(
    Years_Observed   = n(),
    Mean_Discrepancy = mean(Discrepancy, na.rm = TRUE),
    Mean_Rank        = mean(Rank, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  arrange(desc(Mean_Discrepancy))

ignore_voting_admin_vars <- c("VOTED_rc", "VOTEHOW_rc", "VOTEREG_rc")

selected_weighting_vars <- discrepancy_consistency %>%
  filter(!Variable %in% ignore_voting_admin_vars) %>%
  arrange(Mean_Rank) %>%
  slice_head(n = 4)
```

Now, limit the targets to only include the selected weighting variables. 
```{r}
restricted_targets <- lapply(all_targets, function(year_targets) {
  year_targets[names(year_targets) %in% selected_weighting_vars$Variable]
})
```

Created restricted weights and add to df
```{r}
weighting_results_restricted <- run_weighting(c(2006, 2008, 2010, 2012, 2014, 2016, 2018, 2020, 2022), restricted_targets)

CES_06_rc_clean <- add_weights_to_df(CES_06_rc_clean, 2006, weighting_results_restricted, "anesrake_weight_restricted")
CES_08_rc_clean <- add_weights_to_df(CES_08_rc_clean, 2008, weighting_results_restricted, "anesrake_weight_restricted")
CES_10_rc_clean <- add_weights_to_df(CES_10_rc_clean, 2010, weighting_results_restricted, "anesrake_weight_restricted")
CES_12_rc_clean <- add_weights_to_df(CES_12_rc_clean, 2012, weighting_results_restricted, "anesrake_weight_restricted")
CES_14_rc_clean <- add_weights_to_df(CES_14_rc_clean, 2014, weighting_results_restricted, "anesrake_weight_restricted")
CES_16_rc_clean <- add_weights_to_df(CES_16_rc_clean, 2016, weighting_results_restricted, "anesrake_weight_restricted")
CES_18_rc_clean <- add_weights_to_df(CES_18_rc_clean, 2018, weighting_results_restricted, "anesrake_weight_restricted")
CES_20_rc_clean <- add_weights_to_df(CES_20_rc_clean, 2020, weighting_results_restricted, "anesrake_weight_restricted")
CES_22_rc_clean <- add_weights_to_df(CES_22_rc_clean, 2022, weighting_results_restricted, "anesrake_weight_restricted")
update_list_all_years()
```

# COMPUTE CES ERRORS
CES error is measured in the following ways:
  - Candidate choice at the Party-level: compare vote share for the modal party (e.g., Democrats 55%).
  - Candidate choice at the Candidate-level: compare vote share for the modal candidate (e.g., Obama 55%).
  - Demographic and voting variables ("demoplus"): compare true population proportions to CES proportions for demographic and voting behavior variables (e.g., union membership, voter turnout, etc.).

```{r}
# Variables used by the CES for post-stratification in each year
# This is necessary for classifying variables as "primary" or "secondary"
vars_used_ces_weights <- list(
  `2006` = character(0),
  `2008` = c("AGE_GROUP_rc", "EDUC_rc", "SEX_rc", "VOTED_rc"),
  `2010` = c("AGE_GROUP_rc", "EDUC_rc", "SEX_rc", "VOTED_rc"),
  `2012` = c("AGE_GROUP_rc", "EDUC_rc", "SEX_rc", "VOTED_rc"),
  `2014` = c("AGE_GROUP_rc", "EDUC_rc", "SEX_rc", "VOTEREG_rc", "POST_STATE_rc", "GOV_CANDIDATE_rc", "GOV_PARTY_rc", "SENATOR_CANDIDATE_rc", "SENATOR_PARTY_rc"),
  `2016` = c("AGE_GROUP_rc", "EDUC_rc", "SEX_rc", "VOTEREG_rc", "POST_STATE_rc", "GOV_CANDIDATE_rc", "GOV_PARTY_rc", "SENATOR_CANDIDATE_rc", "SENATOR_PARTY_rc"),
  `2018` = c("AGE_GROUP_rc", "EDUC_rc", "SEX_rc", "HISPANIC_rc", "VOTEREG_rc", "POST_STATE_rc", "GOV_CANDIDATE_rc", "GOV_PARTY_rc", "SENATOR_CANDIDATE_rc", "SENATOR_PARTY_rc"),
  `2020` = c("AGE_GROUP_rc", "EDUC_rc", "SEX_rc", "HISPANIC_rc", "VOTEREG_rc", "POST_STATE_rc", "GOV_CANDIDATE_rc", "GOV_PARTY_rc", "SENATOR_CANDIDATE_rc", "SENATOR_PARTY_rc", "PRES_PARTY_rc", "PRES_CANDIDATE_rc"),
  `2022` = c("AGE_GROUP_rc", "EDUC_rc", "SEX_rc", "HISPANIC_rc", "VOTEREG_rc", "POST_STATE_rc", "GOV_CANDIDATE_rc", "GOV_PARTY_rc", "SENATOR_CANDIDATE_rc", "SENATOR_PARTY_rc"
  )
)
```

```{r}
# CES Weighting vars for each year
yearly_ces_wt_vars <- list(
  "2006" = c("v1001", "anesrake_weight_full", "anesrake_weight_restricted"),
  "2008" = c("V201", "anesrake_weight_full", "anesrake_weight_restricted"),
  "2010" = c("V101", "anesrake_weight_full", "anesrake_weight_restricted"),
  "2012" = c("V103", "anesrake_weight_full", "anesrake_weight_restricted"),
  "2014" = c("weight", "anesrake_weight_full", "anesrake_weight_restricted"),
  "2016" = c("commonweight_post", "anesrake_weight_full", "anesrake_weight_restricted"),
  "2018" = c("commonpostweight", "anesrake_weight_full", "anesrake_weight_restricted"),
  "2020" = c("commonpostweight",  "anesrake_weight_full", "anesrake_weight_restricted"),
  "2022" = c("commonpostweight", "anesrake_weight_full", "anesrake_weight_restricted"))
```

```{r}
# Create accuracy configuration object
config <- create_accuracy_config(
  yearly_ces_wt_vars = yearly_ces_wt_vars,
  skip_conditions = skip_conditions,
  NC_flag_df = NC_flag_df,
  vars_used_ces_weights = vars_used_ces_weights,
  
  
  anesrake_full_results       = weighting_results_all,
  anesrake_restricted_results = weighting_results_restricted,
  selected_weighting_vars = selected_weighting_vars
)
```

## Calculate party-level election errors
```{r}
party_errors <- calculate_party_errors(
  year_data_list = list_all_years,
  election_results_df = merged_prop_tables_combined,
  config = config
) %>%
  standardize_variable_names() %>%
  mutate(Specificity = "Party")
```

## Calculate candidate-level election errors
```{r}
# Candidate-level is only calculated at modal level to begin with
candidate_errors_modal <- calculate_candidate_errors(
  year_data_list = list_all_years,
  candidate_returns_df = merged_candidate_level_props,
  config = config
) %>%
  standardize_variable_names() %>%
  mutate(Specificity = "Candidate")

# export to excel sorted by Matching score to review fuzzy matching if needed
# write_excel_csv(candidate_errors_modal, "candidate_errors_modal.csv")
```

## Calculate demovote errors (demographics + voting administration)
```{r}
# Automatically extracts all comparison variables from all_targets
# VOTED_rc (turnout) uses actual state turnout rates as benchmark
# POST_STATE_rc (state residence) uses actual state populations as benchmark
# All other variables use CPS as benchmark

demovote_errors <- calculate_demovote_errors(
  year_data_list = list_all_years,
  all_targets = all_targets,  # Automatically extracts all comparable variables
  config = config,
  turnout_statewide = turnout_statewide,  # For accurate turnout benchmarks
  state_populations = state_populations    # For accurate state sampling benchmarks
) %>%
  standardize_variable_names() %>%
  mutate(Specificity = NA)
```

## National-level errors
The above accuracy functions measure error where the unit of observation is the state. This function treats the country as the unit of observation, so there's just one error per variable per year (x the number of categories for that variable)

```{r}
calculate_demovote_errors_national <- function(year_data_list,
                                               all_targets,
                                               config,
                                               turnout_national = NULL) {

  comparison_vars <- extract_comparison_vars(all_targets)
  all_results <- list()

  for (year_data in year_data_list) {

    year   <- year_data$year
    ces_df <- year_data$CES
    cps_df <- year_data$CPS

    wt_vars <- config$weight_vars[[as.character(year)]]

    for (demo_var in comparison_vars) {

      # ---------------------------------------
      # Skip inherently state-distribution vars
      # ---------------------------------------
      if (demo_var == "POST_STATE_rc") next
      if (!(demo_var %in% names(ces_df))) next

      var_metadata <- determine_variable_metadata(demo_var, year, config)

      # ---------------------------------------
      # Special case: turnout
      # ---------------------------------------
      if (demo_var == "VOTED_rc" && !is.null(turnout_national)) {

        categories <- c("Voted", "Did not vote")
        benchmark_voted <- turnout_national %>%
          filter(Year == year) %>%
          select(-Year) %>%
          unlist(use.names = FALSE)


        benchmark <- c(
          Voted = benchmark_voted,
          `Did not vote` = 1 - benchmark_voted
        )

      } else {
        # ---------------------------------------
        # CPS benchmark variables
        # ---------------------------------------
        if (!(demo_var %in% names(cps_df))) next

        categories <- sort(unique(c(ces_df[[demo_var]], cps_df[[demo_var]])))
        categories <- categories[!is.na(categories)]

        benchmark <- prop.table(
          table(factor(cps_df[[demo_var]], levels = categories))
        )
      }

      # ---------------------------------------
      # CES national estimates
      # ---------------------------------------
      ces_unwt <- prop.table(
        table(factor(ces_df[[demo_var]], levels = categories))
      )

      ces_wt <- weighted_prop_zero(
        ces_df[[demo_var]],
        ces_df[[wt_vars[1]]],
        categories
      )

      anes_full <- weighted_prop_zero(
        ces_df[[demo_var]],
        ces_df$anesrake_weight_full,
        categories
      )

      anes_res <- weighted_prop_zero(
        ces_df[[demo_var]],
        ces_df$anesrake_weight_restricted,
        categories
      )

      # ---------------------------------------
      # Store results
      # ---------------------------------------
      for (cat in categories) {

        all_results[[length(all_results) + 1]] <- tibble(
          Year     = year,
          State    = "NATIONAL",
          Class    = if_else(
            demo_var %in% c("VOTED_rc","VOTEHOW_rc","VOTEREG_rc"),
            "Voting Administration",
            "Demographic"
          ),
          Variable = demo_var,
          Category = cat,

          Variable_Type = var_metadata$variable_type,
          Used_in_ANESRake_Full       = var_metadata$used_in_anesrake_full,
          Used_in_ANESRake_Restricted = var_metadata$used_in_anesrake_restricted,
          Valid_for_Accuracy_Full       = var_metadata$valid_for_accuracy_full,
          Valid_for_Accuracy_Restricted = var_metadata$valid_for_accuracy_restricted,

          Benchmark = benchmark[cat] * 100,

          CES_Unweighted          = ces_unwt[cat] * 100,
          CES_Weighted            = ces_wt[cat] * 100,
          CES_ANESRake_Full       = anes_full[cat] * 100,
          CES_ANESRake_Restricted = anes_res[cat] * 100,

          Error_Unweighted          = (ces_unwt[cat] - benchmark[cat]) * 100,
          Error_CES_Weighted        = (ces_wt[cat]   - benchmark[cat]) * 100,
          Error_ANESRake_Full       = (anes_full[cat] - benchmark[cat]) * 100,
          Error_ANESRake_Restricted = (anes_res[cat]  - benchmark[cat]) * 100,

          n_respondents = sum(!is.na(ces_df[[demo_var]]))
        )
      }
    }
  }

  bind_rows(all_results)
}
```

```{r}
demovote_errors_national <- calculate_demovote_errors_national(
  year_data_list = list_all_years,
  all_targets    = all_targets,
  config         = config,
  turnout_national = turnout_national
) %>%
  standardize_variable_names() %>%
  mutate(Specificity = NA)

demovote_errors_national_modal <- keep_modal_only(demovote_errors_national)
demovote_errors_national_modal_long <- demovote_errors_national_modal %>%
  pivot_errors_long()
```

## Keep modal category only, combine, pivot
```{r}
# Candidate is already modal
party_errors_modal <- keep_modal_only(party_errors)
demovote_errors_modal <- keep_modal_only(demovote_errors)

# Combined
combined_errors <- bind_rows(
  candidate_errors_modal,
  party_errors_modal,
  demovote_errors_modal,
)

# Pivot to long
combined_errors_long <- combined_errors %>%
  pivot_errors_long()
```


# Create Summary Tables
## Metrics base
metrics_base is the core analysis table used for all accuracy evaluations.
It contains every benchmark–estimate comparison in long format, expanded across the three validity regimes that define which variables are appropriate to evaluate.

Each row represents a single state (or district) × variable × weighting method × validity regime comparison, with the associated benchmark, CES estimate, and error.

Concretely, metrics_base is constructed by:
(1) Starting from all error records (candidate, party, demographics),
(2) Duplicating rows across validity regimes where appropriate.

The three validity regimes are:
(1) All — no filtering; all variables are evaluated
(2) ANESRake – Full — only variables valid when ANESRake used the full weighting set
(3) ANESRake – Restricted — only variables valid when ANESRake used the restricted weighting set
```{r}
metrics_base <- bind_rows(

  # ALL (no filtering)
  combined_errors_long %>%
    mutate(Validity_Scheme = "All"),

  # FULL
  combined_errors_long %>%
    filter(Valid_for_Accuracy_Full) %>%
    filter(
      Weighting_Method %in% c(
        "CES-Unweighted",
        "CES-Provided Weights",
        "ANESRake-Full"
      )
    ) %>%
    mutate(Validity_Scheme = "ANESRake - Full"),

  # RESTRICTED
  combined_errors_long %>%
    filter(Valid_for_Accuracy_Restricted) %>%
    filter(
      Weighting_Method %in% c(
        "CES-Unweighted",
        "CES-Provided Weights",
        "ANESRake-Restricted"
      )
    ) %>%
    mutate(Validity_Scheme = "ANESRake - Restricted")
)
```

## Most-Specific
This table uses candidate-level error where available, and otherwise falls back on party-level error.
```{r}
cc_base <- metrics_base %>%
  filter(Class == "Candidate Choice")


cand_keep <- cc_base %>%
  filter(
    Specificity == "Candidate",
    Year != 2006,
    (Variable %in% c("President", "U.S. Senate", "U.S. House") |
      # 2010–: Governor
      (Year >= 2010 & Variable == "Governor")) |
      # 2018–: SecState + AG
      (Year >= 2018 &
         Variable %in% c("Secretary of State","Attorney General")))


party_keep <- cc_base %>%
  filter(
    Specificity == "Party",
    (
      # 2006: all candidate choice from party
      (Year == 2006) |
      # State senator and state rep always party-level
      (Variable %in% c("State Senator", "State Representative")) |
         
      (Year <= 2008 &
         Variable %in% c("Governor")) |

      # 2010–2016: only SecState, AG, State Leg
      (Year <= 2016 &
         Variable %in% c("Secretary of State", "Attorney General"))))

demo_keep <- metrics_base %>%
  filter(Class != "Candidate Choice")

most_specific <- bind_rows(
  demo_keep,
  cand_keep,
  party_keep
) %>%
  arrange(Year, State, Class, Variable, Category)

```

# Export Tables
```{r}
# Metrics base with most_specific_long
write_xlsx(
  metrics_base,
  here("tables_and_figures/input_tables", "metrics_base.xlsx")
)

write_xlsx(
  most_specific,
  here("tables_and_figures/input_tables", "most_specific.xlsx")
)

write_xlsx(
  demovote_errors_national_modal_long,
  here("tables_and_figures/input_tables", "national_errors.xlsx")
)

write_xlsx(
  vars_used_by_anesrake,
  here("tables_and_figures/input_tables", "vars_used_by_anesrake.xlsx")
)
```


## Combining Errors
```{r}
# ===============================
# Party-level + Demovote
# ===============================
party_demovote_wide <- bind_rows(
  party_errors_modal,
  demovote_errors_modal
)

# ===============================
# Candidate-level + Demovote
# ===============================
candidate_demovote_wide <- bind_rows(
  candidate_errors_modal,
  demovote_errors_modal
)

# ===============================
# Most-specific-available table
# ===============================
# Use candidate-level when available, fall back to party-level otherwise

# Define candidate-choice variables
final_cand_choice_vars <- c(
  "President", "U.S. Senate", "U.S. House", "Governor",
  "Secretary of State", "Attorney General", "State Senator", "State Representative"
)

# Candidate-level keepers by year rules
cand_keep <- bind_rows(
  # 2010–2016: candidate-level for Pres/Sen/House/Gov
  candidate_errors_modal %>%
    filter(Class == "Candidate Choice",
           Year >= 2010, Year <= 2016,
           Variable %in% c("President", "U.S. Senate", "U.S. House", "Governor")),

  # 2018–2022: candidate-level adds SecState + AG
  candidate_errors_modal %>%
    filter(Class == "Candidate Choice",
           Year >= 2018, Year <= 2022,
           Variable %in% c("President", "U.S. Senate", "U.S. House", "Governor",
                          "Secretary of State", "Attorney General"))
)

# Party-level fallback by year rules
party_keep <- bind_rows(
  # 2006 & 2008: all candidate-choice from party-level
  party_errors_modal %>%
    filter(Class == "Candidate Choice",
           Year %in% c(2006, 2008),
           Variable %in% final_cand_choice_vars),

  # 2010–2016: party-level for SecState, AG, State Leg
  party_errors_modal %>%
    filter(Class == "Candidate Choice",
           Year >= 2010, Year <= 2016,
           Variable %in% c("Secretary of State", "Attorney General",
                          "State Senator", "State Representative")),

  # 2018–2022: party-level only for State Leg
  party_errors_modal %>%
    filter(Class == "Candidate Choice",
           Year >= 2018, Year <= 2022,
           Variable %in% c("State Senator", "State Representative"))
)

# Most-specific-available: demographics + best available candidate choice
most_specific_wide <- bind_rows(
  demovote_errors_modal,
  cand_keep,
  party_keep
) %>%
  arrange(Year, State, Class, Variable, Category) 

```


## Convert to Long Format
```{r}
party_demovote_long <- pivot_errors_long(party_demovote_wide)
candidate_demovote_long <- pivot_errors_long(candidate_demovote_wide)
most_specific_long <- pivot_errors_long(most_specific_wide)
```

# CREATE SUMMARY TABLES
## Base table for calculations
```{r}
# Create a full table from most_specific_long with a Validity_Scheme flag
#
# Validity_Scheme indicates the regime in which a variable is valid for
# accuracy evaluation:
#
#   All                = all variables (no filtering)
#   ANESRake–Full      = valid for accuracy assuming ANESRake uses the full
#                        set of weighting variables
#   ANESRake–Restricted = valid for accuracy assuming ANESRake uses only the
#                        restricted set of weighting variables

metrics_base <- bind_rows(

  # -----------------------------
  # ALL (no filtering)
  # -----------------------------
  most_specific_long %>%
    mutate(Validity_Scheme = "All"),

  # -----------------------------
  # FULL validity
  # -----------------------------
  most_specific_long %>%
    filter(Valid_for_Accuracy_Full) %>%
    filter(
      Weighting_Method %in% c(
        "CES-Unweighted",
        "CES-Provided Weights",
        "ANESRake-Full"
      )
    ) %>%
    mutate(Validity_Scheme = "ANESRake - Full"),

  # -----------------------------
  # RESTRICTED validity
  # -----------------------------
  most_specific_long %>%
    filter(Valid_for_Accuracy_Restricted) %>%
    filter(
      Weighting_Method %in% c(
        "CES-Unweighted",
        "CES-Provided Weights",
        "ANESRake-Restricted"
      )
    ) %>%
    mutate(Validity_Scheme = "ANESRake - Restricted")
)
```
