---
title: "analysisHQ"
author: "JD Pruett"
date: "2025-04-04"
output: html_document
---

# Environment Setup --------------------------------------
```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = '/Users/jdpruett/Desktop/CES Accuracy Analysis')

# Load functions from other utility scripts
source("recoding_utilities.R")
```

```{r, echo=False, warnings=FALSE, message=FALSE}
# create notin operator
`%notin%` <- Negate(`%in%`)

# Packages & working directory
pckgs <- c("haven", "readxl", "ipumsr", "here", "dplyr", "cdlTools", "anesrake", "writexl", "tidyverse", "rlang", "tibble", "naniar", "weights", "lubridate", "openxlsx", "cpsR")

#For now, this installation is not working
#"zipcodeR"

# This downloads any necessary packages
# if (any(pckgs %notin% rownames(installed.packages())==TRUE)){
# install.packages(pckgs, repos = c(CRAN = "http://cloud.r-project.org"))}

invisible(sapply(pckgs, FUN = require, character.only = TRUE))

# Set working directory.
set_here()
```

# Read-in Data ----------------------------------------------
## CES Datasets
```{r, Read-in CES Data}
CES_06_raw <- read_dta(here("data/CES Data/CES06/cces_2006_common.dta"))
CES_08_raw <- read_dta(here("data/CES Data/CES08/cces_2008_common.dta"))
CES_10_raw <- read_dta(here("data/CES Data/CES10/cces_2010_common_validated.dta"))
CES_12_raw <- read_dta(here("data/CES Data/CES12/commoncontent2012.dta"))
CES_14_raw <- read_dta(here("data/CES Data/CES14/CCES14_Common_Content_Validated.dta"))
CES_16_raw <- read_dta(here("data/CES Data/CES16/CCES16_Common_OUTPUT_Feb2018_VV.dta"))
CES_18_raw <- read_dta(here("data/CES Data/CES18/cces18_common_vv.dta"))
CES_20_raw <- read_dta(here("data/CES Data/CES20/CES20_Common_OUTPUT_vv.dta"))
CES_22_raw <- read_dta(here("data/CES Data/CES22/CES_2022_Common_Output.dta"))
```


## Election Validation Datasets
### National Elections
```{r}
# Governor, U.S. Senate, President
Governor_0622 <- read_excel("data/Election Validation/Final Sets - Clean Fixed/Governor_0622_clean_fixed.xlsx")
Senate_0622 <- read_excel("data/Election Validation/Final Sets - Clean Fixed/Senate_0622_clean_fixed.xlsx")
President_0820 <- read_excel("data/Election Validation/Final Sets - Clean Fixed/[FIN-FIX] President 2008-2020 clean fixed.xlsx")
'/Users/jdpruett/Desktop/CES Accuracy Analysis/data/Election Validation/Final Sets - Clean Fixed/[FIN-FIX] Historical election results data (2006 - 2023) clean fixed.xlsx'

# U.S. House
house_data_path <- "data/Election Validation/Final Sets - Clean Fixed/[FIN-FIX] House 2006-2022 clean fixed.xlsx"
house_sheet_names <- excel_sheets(house_data_path)
house_sheet_names <- house_sheet_names[house_sheet_names != "Sources"]
House_0622 <- house_sheet_names %>%
  lapply(function(sheet) {
    data <- read_excel(path = house_data_path, sheet = sheet)
    data$state <- toupper(data$state)
    return(data)
  }) %>%
  bind_rows()

House_0622 <- House_0622 %>%
  mutate(party_simplified = case_when(
    party_detailed %in% c("DEM", "ADDITIONAL (D)", "ADDITIONAL(D)", "DNL", "DEMOCRATIC-FARMER-LABOR", "DEMOCRATIC") ~ "DEM",
    party_detailed %in% c("ADDITIONAL (R)", "ADDITIONAL(R)") ~ "REP",
    TRUE ~ party_simplified
  ))
```

### State Elections
```{r warning=FALSE}
# State-level elections master file
State_race_master_table <- read_excel("data/Election Validation/Final Sets - Clean Fixed/[FIN-FIX] Historical election results data (2006 - 2023) clean fixed.xlsx")
State_race_master_table$`Election date` <- as.POSIXct(format(State_race_master_table$`Election date`, "%Y-%m-%d"))

# Parsing state-level master file into individual elections
## State senate
State_senate_full <- State_race_master_table[grepl("\\b(senate|senator|SS|SD|KYSENATOR)\\b", State_race_master_table[["Office/district"]], ignore.case = TRUE), ]

# State representative
State_house_full <- State_race_master_table[grepl("\\b(house|assembly|representative|partyresentative|State Rep.|SR|HD|assemblymember|KYREPRESENTATIVE|StateRepresentative|HouseDistrict)\\b|State Rep\\.", State_race_master_table[["Office/district"]], ignore.case = TRUE), ]

# Attorney general
Attorney_general_full <- State_race_master_table[grepl("Attorney", State_race_master_table[["Office/district"]], ignore.case = TRUE), ]

# Secretary of state
Secretary_of_state_full <- State_race_master_table[grepl("Secretary|Secratary", State_race_master_table[["Office/district"]], ignore.case = TRUE), ]
```

### Misc.
```{r}
# CES assigned some respondents to incorrect congressional district in 2020
NC_flag_df <- read.csv("data/misc/NC_responents_incorrect_CD.csv") %>%
  filter(cd_change_flag_post == 1)
```

## CPS Datasets
```{r}
# Set IPUMS API Key (specific to user, free to acquire on IPUMS site)
USER_IPUMS_API_KEY = "59cba10d8a5da536fc06b59dd33fd321a4eb480fa8fe8ad46c9d0d3e"
set_ipums_api_key(USER_IPUMS_API_KEY, save = TRUE, overwrite = FALSE)
```

### Prep monthly sample codes and variables
```{r}
# All
cps_vars_all <- c("WTFINL", "EARNWT", "VOSUPPWT", "STATEFIP", "SEX", "EDUC", "AGE", "CITIZEN", "HISPAN", "OCC", "VETSTAT", "REGION", "EMPSTAT", "CLASSWKR", "UNION", "FAMINC", "VOTERES", "VOREG", "VOTED", "VOWHYNOT", "VOTEHOW")

# 2006
cps_sample_2006 = c('cps2006_01s', 'cps2006_02b', 'cps2006_03b', 'cps2006_04b', 'cps2006_05s', 'cps2006_06s', 'cps2006_07b', 'cps2006_08s', 'cps2006_09s', 'cps2006_10s', 'cps2006_11s', 'cps2006_12s')
# 2008
cps_sample_2008 = c('cps2008_01s', 'cps2008_02b', 'cps2008_03b', 'cps2008_04b', 'cps2008_05s', 'cps2008_06s', 'cps2008_07b', 'cps2008_08s', 'cps2008_09s', 'cps2008_10s', 'cps2008_11s', 'cps2008_12s')
# 2010
cps_sample_2010 = c('cps2010_01s', 'cps2010_02b', 'cps2010_03b', 'cps2010_04b','cps2010_05s', 'cps2010_06s','cps2010_07s', 'cps2010_08s', 'cps2010_09s', 'cps2010_10s', 'cps2010_11s', 'cps2010_12s')
# 2012
cps_sample_2012 = c('cps2012_01s', 'cps2012_02b', 'cps2012_03b', 'cps2012_04b', 'cps2012_05s', 'cps2012_06s', 'cps2012_07s', 'cps2012_08s', 'cps2012_09s', 'cps2012_10s', 'cps2012_11s', 'cps2012_12s')
# 2014
cps_sample_2014 = c('cps2014_01s', 'cps2014_02s', 'cps2014_03b', 'cps2014_04b', 'cps2014_05b', 'cps2014_06s', 'cps2014_07s', 'cps2014_08s', 'cps2014_09s', 'cps2014_10s', 'cps2014_11s', 'cps2014_12s')
# 2016
cps_sample_2016 = c('cps2016_01s', 'cps2016_02s', 'cps2016_03b', 'cps2016_04b', 'cps2016_05b', 'cps2016_06s', 'cps2016_07b', 'cps2016_08s', 'cps2016_09s', 'cps2016_10s', 'cps2016_11s', 'cps2016_12s')
# 2018
cps_sample_2018 = c('cps2018_01s', 'cps2018_02s', 'cps2018_03b', 'cps2018_04b', 'cps2018_05s', 'cps2018_06s', 'cps2018_07s', 'cps2018_08s', 'cps2018_09s', 'cps2018_10s', 'cps2018_11s', 'cps2018_12s')
# 2020
cps_sample_2020 = c('cps2020_01s', 'cps2020_02s', 'cps2020_03b', 'cps2020_04b', 'cps2020_05b', 'cps2020_06s', 'cps2020_07b', 'cps2020_08s', 'cps2020_09b', 'cps2020_10s', 'cps2020_11s', 'cps2020_12s')
# 2022
cps_sample_2022 = c('cps2022_01s', 'cps2022_02s', 'cps2022_03b', 'cps2022_04s', 'cps2022_05s', 'cps2022_06s', 'cps2022_07s', 'cps2022_08s', 'cps2022_09s', 'cps2022_10s', 'cps2022_11s', 'cps2022_12s')

```

### Read in CPS datasets to global environment
```{r}
# All years
extract_codes <- list()

for (year in seq(2006, 2022, by = 2)) {
  suffix <- substr(year, 3, 4)
  extract_name <- paste("CPS Extract for", year)  # Dynamic extract name
  cps_sample_var <- paste("cps_sample", year, sep = "_")  # Dynamic sample variable name
  df_name <- paste("CPS", suffix, "raw", sep = "_")

  
  # Dynamically create the extraction code using the variables
  extract_code <- paste0(
    "  define_extract_cps(\n",
    "    \"", extract_name, "\",\n",
    "    ", cps_sample_var, ",\n",
    "    cps_vars_all) %>%\n",
    "  submit_extract() %>%\n",
    "  wait_for_extract() %>%\n",
    "  download_extract() %>%\n",
    "  read_ipums_micro()"
  )
  
  extract_codes[[df_name]] <- extract_code
}

# Evaluate the constructed extraction codes in the global environment
for (df_name in names(extract_codes)) {
  assign(df_name, eval(parse(text = extract_codes[[df_name]])), envir = .GlobalEnv)
}
```

## Subset datasets --------------
### Subset CES to only include relevant variables
```{r}
CES_06_subset_vars <- c(
  "v1000", "v1001", "v1002", "v1003", "v1004", "v2004", "v2018", "v2020", "v3081", "v3083", "v1006", "v2030", "v2073", "v2082", "v2033", "v2032", "v3004", "v4004", "v4006", "v4005", "v4014", "v4015", "v4013", "v4017", "v4018", "v4019", "v4020", "v4021", "v5009", "v5010", "v5011", "v5012", "v5005", "v5006", "v5007", "v5008", "v5001", "v5002", "v5003", "v5004")

CES_08_subset_vars <- c(
  "V100", "V201", "V206", "V207", "V208", "V250", "V251", "V202", "V205", "V204", "V213", "V252", "V253", "V258", "V257", "V247", "V259", "V264", "V501", "V502", "V553", "V555", "V601", "V602", "V603", "V612", "V653", "V655", "V657", "V672", "CC328_1", "CC328_3", "CC334", "CC410", "CC411", "CC411b", "CC412", "CC413", "CC414_2", "CC414_1", "V209", "CC329", "CC333", "V246", "V203", "CC403", "CC405", "CC404", "CC326b", "CC337", "CC337S2", "CC340", "CC338", "V618", "V619")

CES_10_subset_vars <- c(
  "V100", "V101", "V202_post", "V204_post", "V205_post", "V206", "V206_post", "V207", "V208", "V213", "V252", "V254", "V276", "V276_post", "V564_post", "V567_post", "V548_post", "V551_post", "V533_post", "V536_post", "V554_post", "V556_post", "V570_post", "V572_post","V574_post", "V576_post", "CC351", "V209", "V264", "V250", "V246", "V290", "CC401", "V203", "CC403", "CC402a", "CC410a", "CC410b", "CC412", "CC411", "CC413a", "CC413b", "CC413c", "CC413d")

CES_12_subset_vars <- c(
  "V101", "V103", "inputzip", "votereg_f", "regzip", "inputstate", "inputstate_post", "cdid113", "cdid113_post", "gender", "educ", "birthyr", "milstat_1", "milstat_3", "CC351", "votereg", "employ", "employercat", "union", "ownhome", "faminc",
  "hispanic", "CC401", "CC402a", "CC403", "CC410a", "CC410b", "CC355", "CC412", "CC411", "CC413a", "CC413b", "CC413c", "CC413d", "GovCand1Name_post", "GovCand1Party_post", "GovCand2Name_post", "GovCand2Party_post", "SenCand1Name_post", "SenCand1Party_post", "SenCand2Name_post", "SenCand2Party_post", "SenCand3Name_post", "SenCand3Party_post", "HouseCand1Name_post", "HouseCand1Party_post", "HouseCand2Name_post", "HouseCand2Party_post", "HouseCand3Name_post", "HouseCand3Party_post")

CES_14_subset_vars <- c(
  "V101", "weight", "inputzip_post", "votereg_f_post", "regzip_post", "inputstate", "inputstate_post", "cdid", "cdid_post", "gender", "educ", "birthyr", "milstat_1", "milstat_3", "CC351", 
  "votereg", "employ", "union", "ownhome", "faminc",
  "hispanic", "CC401", "CC402a", "CC403", "CC410b", "CC355x", "CC412", "CC411", "CC413a", "CC413b", "CC413c", "CC413d", "GovCand1Name_post", "GovCand1Party_post", "GovCand2Name_post", "GovCand2Party_post", "SenCand1Name_post", "SenCand1Party_post", "SenCand2Name_post", "SenCand2Party_post", "SenCand3Name_post", "SenCand3Party_post", "HouseCand1Name_post", "HouseCand1Party_post", "HouseCand2Name_post", "HouseCand2Party_post", "HouseCand3Name_post", "HouseCand3Party_post")

CES_16_subset_vars <- c(
  "V101", "inputzip_post", "votereg_f_post", "regzip_post", "commonweight_vv", "commonweight_vv_post", "commonweight_vv_lgbt", "commonweight", "commonweight_post", "inputstate", "inputstate_post", "cdid115", "cdid115_post", "gender", "educ", "birthyr", "milstat_1", "milstat_3", "CC16_361", "votereg", "employ", "union", "ownhome", "faminc", "hispanic", "CC16_401", "CC16_402a", "CC16_410a", "CC16_403", "CC16_410b", "CC16_412", "CC16_411", "CC16_413a", "CC16_413b", "CC16_413c", "CC16_413d", "GovCand1Name_post", "GovCand1Party_post", "GovCand2Name_post", "GovCand2Party_post", "GovCand3Name_post", "GovCand3Party_post", "SenCand1Name_post", "SenCand1Party_post", "SenCand2Name_post", "SenCand2Party_post", "SenCand3Name_post", "SenCand3Party_post", "SenCand4Name_post", "SenCand4Party_post", "HouseCand1Name_post", "HouseCand1Party_post", "HouseCand2Name_post", "HouseCand2Party_post", "HouseCand3Name_post", "HouseCand3Party_post", "HouseCand4Name_post", "HouseCand4Party_post", "HouseCand5Name_post", "HouseCand5Party_post", "HouseCand6Name_post", "HouseCand6Party_post", "HouseCand7Name_post", "HouseCand7Party_post", "HouseCand8Name_post", "HouseCand8Party_post", "HouseCand9Name_post", "HouseCand9Party_post", "HouseCand10Name_post", "HouseCand10Party_post", "HouseCand11Name_post", "HouseCand11Party_post")

CES_18_subset_vars <- c(
  "caseid", "inputzip_post", "votereg_f_post", "regzip_post", "commonweight", "commonpostweight", "vvweight", "vvweight_post", "inputstate", "inputstate_post", "cdid116", "cdid116_post", "gender", "educ", "birthyr", "cit1", "milstat_1", "milstat_3", "region", "votereg", "employ", "union", "ownhome", "faminc_new", "hispanic", "CC18_401", "CC18_402a", "CC18_403", "CC18_410b", "CC18_410bx", "CC18_412", "CC18_353x", "CC18_411", "CC18_352x", "CC18_420a", "CC18_420b", "CC18_420c", "CC18_413c", "CC18_413d", "GovCand1Name_post", "GovCand1Party_post", "GovCand2Name_post", "GovCand2Party_post", "GovCand3Name_post", "GovCand3Party_post", "SenCand1Name_post", "SenCand1Party_post", "SenCand2Name_post", "SenCand2Party_post", "SenCand3Name_post", "SenCand3Party_post", "AttCand1Name", "AttCand1Party", "AttCand2Name", "AttCand2Party", "SecCand1Name", "SecCand1Party", "SecCand2Name", "SecCand2Party", "HouseCand1Name_post", "HouseCand1Party_post", "HouseCand2Name_post", "HouseCand2Party_post", "HouseCand3Name_post", "HouseCand3Party_post")

CES_20_subset_vars <- c(
  "caseid", "inputzip_post", "votereg_f_post", "regzip_post", "commonweight", "commonpostweight", "vvweight", "vvweight_post", "inputstate", "inputstate_post", "cdid116", "cdid116_post", "gender", "educ", "birthyr", "cit1", "milstat_1", "milstat_3", "CC20_361", "region", "votereg", "employ", "union", "ownhome", "faminc_new", "hispanic", "CC20_401", "CC20_402a", "CC20_403", "CC20_410", "CC20_411", "CC20_365_voted", "CC20_411b", "CC20_365b_voted", "CC20_412", "CC20_413", "CC20_414a", "CC20_414b", "CC20_415c", "CC20_415d", "GovCand1Name_post", "GovCand1Party_post", "GovCand2Name_post", "GovCand2Party_post", "SenCand1Name_post", "SenCand1Party_post", "SenCand2Name_post", "SenCand2Party_post", "AttCand1Name", "AttCand1Party", "AttCand2Name", "AttCand2Party", "AttCand3Name", "AttCand3Party", "SecCand1Name", "SecCand1Party", "SecCand2Name", "SecCand2Party", "HouseCand1Name_post", "HouseCand1Party_post", "HouseCand2Name_post", "HouseCand2Party_post", "HouseCand3Name_post", "HouseCand3Party_post", "HouseCand4Name_post", "HouseCand4Party_post", "HouseCand5Name_post", "HouseCand5Party_post", "HouseCand6Name_post", "HouseCand6Party_post", "HouseCand7Name_post", "HouseCand7Party_post", "HouseCand8Name_post", "HouseCand8Party_post", "HouseCand9Name_post", "HouseCand9Party_post")

CES_22_subset_vars <- c(
"caseid", "inputzip_post", "votereg_f_post", "regzip_post", "commonweight", "commonpostweight", "inputstate", "inputstate_post", "cdid118", "cdid118_post", "gender4", "educ", "birthyr", "cit1", "milstat_1", "milstat_3", "CC22_361", "region", "votereg", "employ", "union", "ownhome", "faminc_new", "hispanic", "CC22_401", "CC22_402a", "CC22_403", "CC22_411", "CC22_365_voted", "CC22_411b", "CC22_365b_voted", "CC22_412", "CC22_413", "CC22_414a", "CC22_414b", "CC22_415c", "CC22_415d", "GovCand1Name_post", "GovCand1Party_post", "GovCand2Name_post", "GovCand2Party_post", "GovCand3Name_post", "GovCand3Party_post", "SenCand1Name_post", "SenCand1Party_post", "SenCand2Name_post", "SenCand2Party_post", "SenCand3Name_post", "SenCand3Party_post", "SenCand4Name_post", "SenCand4Party_post", "AttCand1Name", "AttCand1Party", "AttCand2Name", "AttCand2Party", "AttCand3Name", "AttCand3Party", "SecCand1Name", "SecCand1Party", "SecCand2Name", "SecCand2Party", "SecCand3Name", "SecCand3Party", "HouseCand1Name_post", "HouseCand1Party_post", "HouseCand2Name_post", "HouseCand2Party_post", "HouseCand3Name_post", "HouseCand3Party_post", "HouseCand4Name_post", "HouseCand4Party_post", "HouseCand5Name_post", "HouseCand5Party_post", "HouseCand6Name_post", "HouseCand6Party_post", "HouseCand7Name_post", "HouseCand7Party_post", "HouseCand8Name_post", "HouseCand8Party_post")

CES_06_raw_subset <- CES_06_raw[, CES_06_subset_vars]
CES_08_raw_subset <- CES_08_raw[, CES_08_subset_vars]
CES_10_raw_subset <- CES_10_raw[, CES_10_subset_vars]
CES_12_raw_subset <- CES_12_raw[, CES_12_subset_vars]
CES_14_raw_subset <- CES_14_raw[, CES_14_subset_vars]
CES_16_raw_subset <- CES_16_raw[, CES_16_subset_vars]
CES_18_raw_subset <- CES_18_raw[, CES_18_subset_vars]
CES_20_raw_subset <- CES_20_raw[, CES_20_subset_vars]
CES_22_raw_subset <- CES_22_raw[, CES_22_subset_vars]
```

### Subset CPS to exclude respondents below 18 (CES is limited to 18+)
```{r}
# Define the years range
years <- seq(2006, 2022, by = 2)

# Iterate over the years
for (year in years) {
  # Extract the last two digits of the year
  year_short <- substr(year, 3, 4)
  
  # Define the variable names for the current year
  raw_df_name <- paste0("CPS_", year_short, "_raw")
  subset_vars <- intersect(colnames(get(raw_df_name)), cps_vars_all)

  # Subset the raw dataframe using the filtered variable list
  subset_df <- get(raw_df_name)[, subset_vars, drop = FALSE]
  
  # print any vars in cps_vars_all that are not in subset_vars
  print(paste("Missing vars for CES", year_short))
  print(setdiff(cps_vars_all, subset_vars))
  
  # Reorder the columns to match the order in cps_vars_all
  subset_df <- subset_df[, match(cps_vars_all, colnames(subset_df))]
  
  # Filter under 18
  subset_df <- subset_df %>%
    filter(AGE >= 18)
  
  # Assign the subset dataframe back to the global environment
  assign(paste0(raw_df_name, "_subset"), subset_df, envir = .GlobalEnv)
}

```

### Update CES/CPS data structures
```{r}
# list_all_years contains a list of all CES and CPS datasets, organized by year. As changes are made to these datasets, update_list_all_years is called throughout the script, ensuring that subsequent analysis is working on the most updated versions
update_list_all_years <- function() {
  list_all_years <- list(
  list(CPS = CPS_06_rc_clean, CES = CES_06_rc_clean, year = 2006),
  list(CPS = CPS_08_rc_clean, CES = CES_08_rc_clean, year = 2008),
  list(CPS = CPS_10_rc_clean, CES = CES_10_rc_clean, year = 2010),
  list(CPS = CPS_12_rc_clean, CES = CES_12_rc_clean, year = 2012),
  list(CPS = CPS_14_rc_clean, CES = CES_14_rc_clean, year = 2014),
  list(CPS = CPS_16_rc_clean, CES = CES_16_rc_clean, year = 2016),
  list(CPS = CPS_18_rc_clean, CES = CES_18_rc_clean, year = 2018),
  list(CPS = CPS_20_rc_clean, CES = CES_20_rc_clean, year = 2020),
  list(CPS = CPS_22_rc_clean, CES = CES_22_rc_clean, year = 2022)
  )
  
  assign("list_all_years", list_all_years, envir = .GlobalEnv)
}

update_ces_rc_datasets <- function() {
  ces_rc_datasets <- setNames(
  list(CES_06_rc, CES_08_rc, CES_10_rc, CES_12_rc, CES_14_rc, CES_16_rc, CES_18_rc, CES_20_rc, CES_22_rc),
  c("2006", "2008", "2010", "2012", "2014", "2016", "2018", "2020", "2022")
)
  assign("ces_rc_datasets", ces_rc_datasets, envir = .GlobalEnv)
}
```

# Recoding CES and CPS
## Primary recoding
```{r message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE, results='hide'}
# CES and CPS are recoded to have corresponding levels for comparable variables. Ensure that CODEBOOK.xlsm is in the code directory.
apply_codebook_and_recode(6)
apply_codebook_and_recode(8)
apply_codebook_and_recode(10)
apply_codebook_and_recode(12)
apply_codebook_and_recode(14)
apply_codebook_and_recode(16)
apply_codebook_and_recode(18)
apply_codebook_and_recode(20)
apply_codebook_and_recode(22)
```

## Additional recoding
### Location structures
```{r}
pre_post_location_vars <- list(
  # Pre and Post same because non provided
  "2006" = list(
    pre_state = "v1002",
    post_state = "v1002"
  ),
  "2008" = list(
    pre_state = "V206",
    post_state = "V259"
  ),
  "2010" = list(
    pre_state = "V206",
    post_state = "V206_post"
  ),
  "2012" = list(
    pre_state = "inputstate",
    post_state = "inputstate_post"
  ),
  "2014" = list(
    pre_state = "inputstate",
    post_state = "inputstate_post"
  ),
  "2016" = list(
    pre_state = "inputstate",
    post_state = "inputstate_post"
  ),
  "2018" = list(
    pre_state = "inputstate",
    post_state = "inputstate_post"
  ),
  "2020" = list(
    pre_state = "inputstate",
    post_state = "inputstate_post"
  ),
  "2022" = list(
    pre_state = "inputstate",
    post_state = "inputstate_post"
  )
)

make_state_to_region <- function() {
  region_mapping <- list(
    "Northeast" = c("CT", "ME", "MA", "NH", "RI", "VT", "NJ", "NY", "PA"),
    "Midwest" = c("IL", "IN", "MI", "OH", "WI", "IA", "KS", "MN", "MO", "NE", "ND", "SD"),
    "South" = c("DE", "DC", "FL", "GA", "MD", "NC", "SC", "VA", "WV", "AL", "KY", "MS", "TN", "AR", "LA", "OK", "TX"),
    "West" = c("AZ", "CO", "ID", "MT", "NV", "NM", "UT", "WY", "AK", "CA", "HI", "OR", "WA"))

  # Flatten the mapping into a named vector
  state_abbr <- unlist(region_mapping)
  regions <- rep(names(region_mapping), times = lengths(region_mapping))
  setNames(regions, state_abbr)
}

state_to_region <- make_state_to_region()
```

### Manual recodes
```{r warning=FALSE}
year_mapping <- c("2006" = 1, "2008" = 2, "2010" = 3, "2012" = 4, "2014" = 5, "2016" = 6, "2018" = 7, "2020" = 8, "2022" = 9)

# Apply age groups to CES data frames
CES_06_rc$AGE_GROUP_rc <- calculate_age_groups(2006 - CES_06_rc$v2020)
CES_08_rc$AGE_GROUP_rc <- calculate_age_groups(2008 - CES_08_rc$V207)
CES_10_rc$AGE_GROUP_rc <- calculate_age_groups(2010 - CES_10_rc$V207)
CES_12_rc$AGE_GROUP_rc <- calculate_age_groups(2012 - CES_12_rc$birthyr)
CES_14_rc$AGE_GROUP_rc <- calculate_age_groups(2014 - CES_14_rc$birthyr)
CES_16_rc$AGE_GROUP_rc <- calculate_age_groups(2016 - CES_16_rc$birthyr)
CES_18_rc$AGE_GROUP_rc <- calculate_age_groups(2018 - CES_18_rc$birthyr)
CES_20_rc$AGE_GROUP_rc <- calculate_age_groups(2020 - CES_20_rc$birthyr)
CES_22_rc$AGE_GROUP_rc <- calculate_age_groups(2022 - CES_22_rc$birthyr)

# Apply age groups to CPS data frames
CPS_06_rc$AGE_GROUP_rc <- calculate_age_groups(CPS_06_rc$AGE)
CPS_08_rc$AGE_GROUP_rc <- calculate_age_groups(CPS_08_rc$AGE)
CPS_10_rc$AGE_GROUP_rc <- calculate_age_groups(CPS_10_rc$AGE)
CPS_12_rc$AGE_GROUP_rc <- calculate_age_groups(CPS_12_rc$AGE)
CPS_14_rc$AGE_GROUP_rc <- calculate_age_groups(CPS_14_rc$AGE)
CPS_16_rc$AGE_GROUP_rc <- calculate_age_groups(CPS_16_rc$AGE)
CPS_18_rc$AGE_GROUP_rc <- calculate_age_groups(CPS_18_rc$AGE)
CPS_20_rc$AGE_GROUP_rc <- calculate_age_groups(CPS_20_rc$AGE)
CPS_22_rc$AGE_GROUP_rc <- calculate_age_groups(CPS_22_rc$AGE)


# Region
CES_06_rc$REGION_rc <- as.factor(state_to_region[CES_06_rc$v1002])
CES_08_rc$REGION_rc <- as.factor(state_to_region[fips(CES_08_rc$V206, to = "abbreviation")])
CES_10_rc$REGION_rc <- as.factor(state_to_region[fips(CES_10_rc$V206, to = "abbreviation")])
CES_12_rc$REGION_rc <- as.factor(state_to_region[fips(CES_12_rc$inputstate, to = "abbreviation")])
CES_14_rc$REGION_rc <- as.factor(state_to_region[fips(CES_14_rc$inputstate, to = "abbreviation")])
CES_16_rc$REGION_rc <- as.factor(state_to_region[fips(CES_16_rc$inputstate, to = "abbreviation")])
CES_18_rc$REGION_rc <- as.factor(state_to_region[fips(CES_18_rc$inputstate, to = "abbreviation")])
CES_20_rc$REGION_rc <- as.factor(state_to_region[fips(CES_20_rc$inputstate, to = "abbreviation")])
CES_22_rc$REGION_rc <- as.factor(state_to_region[fips(CES_22_rc$inputstate, to = "abbreviation")])

# Add pre and post state
update_ces_rc_datasets()
add_location_vars(list_all_years, pre_post_location_vars)
update_ces_rc_datasets()


# Military Status in CPS
cps_rc_datasets <- list(CPS_06_rc, CPS_08_rc, CPS_10_rc, CPS_12_rc, CPS_14_rc, CPS_16_rc, CPS_18_rc, CPS_20_rc, CPS_22_rc)

for (i in seq_along(cps_rc_datasets)) {
  cps_rc_datasets[[i]]$CUR_MIL_rc <- as.factor(ifelse(
    cps_rc_datasets[[i]]$EMPSTAT == 1,
    "Serving",
    ifelse(cps_rc_datasets[[i]]$EMPSTAT == 0, NA, "Not serving")
  ))
  print(paste("CPS Current Military Done", names(year_mapping)[i]))
}

list2env(setNames(cps_rc_datasets, c("CPS_06_rc", "CPS_08_rc", "CPS_10_rc", "CPS_12_rc", "CPS_14_rc", "CPS_16_rc", "CPS_18_rc", "CPS_20_rc", "CPS_22_rc")), envir = .GlobalEnv)

party_mapping <- c("1" = "Democratic", "2" = "Republican", "3" = "Independent")

#In 2006, it always defaulted that first candidate was Democrat and second was Republican
CES_06_rc <- CES_06_rc %>%
  mutate(
    # Governor
    v5010 = party_mapping[as.character(v5010)],
    v5012 = party_mapping[as.character(v5012)],
    # US Senate
    v5006 = party_mapping[as.character(v5006)],
    v5008 = party_mapping[as.character(v5008)]
  )



## START HERE
# Adding House_rc 
update_ces_rc_datasets()
# Loop through each year and recode HOUSE_rc based on the candidate names available for that year
for (year in names(ces_rc_datasets)) {
  if (year %in% c("2006", "2008", "2010")) next
  
  # Recode the data
  recoded_data <- recode_house_rc(ces_rc_datasets[[year]], yrly_ces_house_candidate_names[[year]])
  
  # Assign the recoded data to its original name in the global environment
  assign(ces_rc_df_name_mapping[year], recoded_data, envir = .GlobalEnv)
}

# Add President names and party columns
CES_08_rc$PRES_DEM_CAND <- "OBAMA"
CES_08_rc$PRES_DEM_PARTY <- "Democratic"
CES_08_rc$PRES_REP_CAND <- "MCCAIN"
CES_08_rc$PRES_REP_PARTY <- "Republican"

CES_12_rc$PRES_DEM_CAND <- "OBAMA"
CES_12_rc$PRES_DEM_PARTY <- "Democratic"
CES_12_rc$PRES_REP_CAND <- "ROMNEY"
CES_12_rc$PRES_REP_PARTY <- "Republican"

CES_16_rc$PRES_DEM_CAND <- "CLINTON"
CES_16_rc$PRES_DEM_PARTY <- "Democratic"
CES_16_rc$PRES_REP_CAND <- "TRUMP"
CES_16_rc$PRES_REP_PARTY <- "Republican"

CES_20_rc$PRES_DEM_CAND <- "BIDEN"
CES_20_rc$PRES_DEM_PARTY <- "Democratic"
CES_20_rc$PRES_REP_CAND <- "TRUMP"
CES_20_rc$PRES_REP_PARTY <- "Republican"

# Clean up NAs
CES_06_rc_clean <- clean_NAs(CES_06_rc)
CPS_06_rc_clean <- clean_NAs(CPS_06_rc)

CES_08_rc_clean <- clean_NAs(CES_08_rc)
CPS_08_rc_clean <- clean_NAs(CPS_08_rc)

CES_10_rc_clean <- clean_NAs(CES_10_rc)
CPS_10_rc_clean <- clean_NAs(CPS_10_rc)

CES_12_rc_clean <- clean_NAs(CES_12_rc)
CPS_12_rc_clean <- clean_NAs(CPS_12_rc)

CES_14_rc_clean <- clean_NAs(CES_14_rc)
CPS_14_rc_clean <- clean_NAs(CPS_14_rc)

CES_16_rc_clean <- clean_NAs(CES_16_rc)
CPS_16_rc_clean <- clean_NAs(CPS_16_rc)

CES_18_rc_clean <- clean_NAs(CES_18_rc)
CPS_18_rc_clean <- clean_NAs(CPS_18_rc)

CES_20_rc_clean <- clean_NAs(CES_20_rc)
CPS_20_rc_clean <- clean_NAs(CPS_20_rc)

CES_22_rc_clean <- clean_NAs(CES_22_rc)
CPS_22_rc_clean <- clean_NAs(CPS_22_rc)

CES_06_rc_clean <- as.data.frame(CES_06_rc_clean)
CES_08_rc_clean <- as.data.frame(CES_08_rc_clean)
CES_10_rc_clean <- as.data.frame(CES_10_rc_clean)
CES_12_rc_clean <- as.data.frame(CES_12_rc_clean)
CES_14_rc_clean <- as.data.frame(CES_14_rc_clean)
CES_16_rc_clean <- as.data.frame(CES_16_rc_clean)
CES_18_rc_clean <- as.data.frame(CES_18_rc_clean)
CES_20_rc_clean <- as.data.frame(CES_20_rc_clean)
CES_22_rc_clean <- as.data.frame(CES_22_rc_clean)

update_list_all_years()
```

